---
title: "Boxplots total AMR and procrustes rotation comparing AMR with Tax"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
Creating figure 5 containing bloxplots of total AMR and procrustes rotation comparing AMR patterns with taxonomic patterns in unspiked. SF14 the same as Fig. 5 but spiked samples  

### Packages
```{r}
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2")  
library(stringr) #Simple, consistent wrappers for common string operations. Like the str_replace function. help(package="stringr")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr")
library(gridExtra)
library(grid)
library(lattice)
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
library(reshape2)
library(tidyr)
```

## Analysis Total AMR
### 1 Read in data
```{r}
AMRFPKM <- read.delim(file="AMRFPKMBac20180919.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```


### 2 Subsetting and create list with boxplots unspiked
```{r}
###1 Define variables
######Define variables used in the analysis

#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<-c("P1unspiked", "P2unspiked", "S1unspiked", "S2unspiked") #All, Allspiked, Allunspiked, PF, SW, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both


BoxplotListTotal <- list()

for (i in Subset) {
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting metadata
if (i=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (i=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (i=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (i=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (i=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (i=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (i=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (i=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (i=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (i=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (i=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (i=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (i=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (i=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (i=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (i=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (i=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
AMRFPKM2<-select(AMRFPKM, one_of(Metadata2$Sample))


###Wanted to plot total AMR per sample
Metadata2$TotalAMR<-colSums(AMRFPKM2)

#Change order on x-axis
Metadata2$Replicate_Boxplot<-as.factor(Metadata2$Replicate_Boxplot)
levels(Metadata2$Replicate_Boxplot)
Metadata2$Replicate_Boxplot<-ordered(Metadata2$Replicate_Boxplot, levels=c("0h", "16h-80C", "64h-80C", "16h-20C", "64h-20C", "16h5C", "64h5C", "16h22C", "64h22C"))

#Reads_Total, Reads_AfterTrim, Mapped, PercentUnmapped. Change 3 places
BoxplotListTotal[[i]] <- ggplot(Metadata2, aes_string(x="Replicate_Boxplot", y="TotalAMR", group="Replicate_Boxplot")) +
  geom_boxplot(aes(fill=Replicate_Boxplot)) + 
  facet_grid(SpikedUnspiked ~ Sample_type_new, scales="free") + 
  #ggtitle(paste("Genus", i, SubExp, sep=" ")) + 
  #xlab("Storage conditions") + 
  ylab("Total AMR per Sample [FPKM]") + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none", strip.text.y = element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank()) + 
  scale_fill_manual(values=c("#999999", "#0571b0", "#0571b0", "#92c5de", "#92c5de", "#f4a582", "#f4a582", "#ca0020", "#ca0020")) #+ 
  #scale_y_continuous(breaks = seq(20, 100, by = 5))
}

#Have the plots stored in lists
lay <- rbind(c(1,2,3,4))
pdf(paste("TotalAMRunspiked.pdf", sep=""), width=8.5, height=2.5)
grid.arrange(BoxplotListTotal$P1unspiked, BoxplotListTotal$P2unspiked, BoxplotListTotal$S1unspiked, BoxplotListTotal$S2unspiked, layout_matrix = lay)  
dev.off() 
```

### 3 Subsetting and create list with boxplots spiked
```{r}
###1 Define variables
######Define variables used in the analysis

#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<-c("P1spiked", "P2spiked", "S1spiked", "S2spiked") #All, Allspiked, Allunspiked, PF, SW, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

for (i in Subset) {
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting metadata
if (i=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (i=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (i=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (i=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (i=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (i=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (i=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (i=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (i=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (i=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (i=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (i=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (i=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (i=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (i=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (i=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (i=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
AMRFPKM2<-select(AMRFPKM, one_of(Metadata2$Sample))


###Wanted to plot total AMR per sample
Metadata2$TotalAMR<-colSums(AMRFPKM2)

#Change order on x-axis
Metadata2$Replicate_Boxplot<-as.factor(Metadata2$Replicate_Boxplot)
levels(Metadata2$Replicate_Boxplot)
Metadata2$Replicate_Boxplot<-ordered(Metadata2$Replicate_Boxplot, levels=c("0h", "16h-80C", "64h-80C", "16h-20C", "64h-20C", "16h5C", "64h5C", "16h22C", "64h22C"))

#Reads_Total, Reads_AfterTrim, Mapped, PercentUnmapped. Change 3 places
BoxplotListTotal[[i]] <- ggplot(Metadata2, aes_string(x="Replicate_Boxplot", y="TotalAMR", group="Replicate_Boxplot")) +
  geom_boxplot(aes(fill=Replicate_Boxplot)) + 
  facet_grid(SpikedUnspiked ~ Sample_type_new, scales="free") + 
  #ggtitle(paste("Genus", i, SubExp, sep=" ")) + 
  #xlab("Storage conditions") + 
  ylab("Total AMR per Sample [FPKM]") + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none", strip.text.y = element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank()) + 
  scale_fill_manual(values=c("#999999", "#0571b0", "#0571b0", "#92c5de", "#92c5de", "#f4a582", "#f4a582", "#ca0020", "#ca0020")) #+ 
  #scale_y_continuous(breaks = seq(20, 100, by = 5))
}

#Have the plots stored in lists
lay <- rbind(c(1,2,3,4))
pdf(paste("TotalAMRspiked.pdf", sep=""), width=8.5, height=2.5)
grid.arrange(BoxplotListTotal$P1spiked, BoxplotListTotal$P2spiked, BoxplotListTotal$S1spiked, BoxplotListTotal$S2spiked, layout_matrix = lay)  
dev.off() 
```

## Analysis Total AMR
### 1 Read in additional data
```{r}
Tax <- read.delim(file="Taxonomy20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```

### 2 Subsetting and create list with procrustes plots
```{r}
#Define variables used in the analysis 
Subset<- c("P1spiked", "P1unspiked", "P2spiked", "P2unspiked", "S1spiked", "S1unspiked", "S2spiked", "S2unspiked") #All, Allspiked, Allunspiked, PF, PFspiked, PFunspiked, SW, SWspiked, SWunspiked, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Filtering which organisms to include
OrgFlt<-"All" #All, Bac, BacSacCry, BacArcVirSacCry 

Stand<-"hellinger" #total, max, freq, normalize, range, standardize, pa, chi.square, hellinger, log, DESeqTrans

#Vegdist dissimilarity method
Dist<-"bray" #manhattan, euclidean, canberra, bray, kulczynski, jaccard, gower, altGower, morisita, horn, mountford, raup, binomial, chao, cao or mahalanobis

#Define coloring schemes
Storagecol<-c("Direct" = "#999999", "-80" = "#0571b0", "-20" = "#92c5de", "5" = "#f4a582", "22" = "#ca0020")
Experimentcol<-c("Pig_feces_1" = "#663300", "Pig_feces_2" = "#FF9900", "Sewage_1" = "#006600", "Sewage_2" = "#33FF33")

## Create a list to hold the plot objects.
ProcrustesList <- list()
#ProerrorsList <- list()


for (i in Subset) {
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting Metadata2
if (i=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (i=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (i=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (i=="PFspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked" | Sample_type == "Pig_feces_2_spiked")
} else if (i=="PFunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1" | Sample_type == "Pig_feces_2")
} else if (i=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (i=="SWspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked" | Sample_type == "Sewage_2_spiked")
} else if (i=="SWunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1" | Sample_type == "Sewage_2")
} else if (i=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (i=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (i=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (i=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (i=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (i=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (i=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (i=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (i=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (i=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (i=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (i=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (i=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
AMRFPKM2<-select(AMRFPKM, one_of(Metadata2$Sample))
Tax2<-select(Tax, one_of(Metadata2$Sample))

#Make protest
prot<-protest(cmdscale(vegdist(decostand(t(Tax2), method="hellinger"), method="bray")), cmdscale(vegdist(decostand(t(AMRFPKM2), method="hellinger"), method="bray")))
prot
#summary(prot)
plot(prot)
plot(prot, kind=2)

Title <- paste("Procrustes", i)
Subtitle <- paste("Correlation = ", round(prot$t0, digits = 4), " (SS = ", round(prot$ss, digits = 4), "), p-value = ", round(prot$signif, digits = 4), sep="") 

#Extract procrustes info
plot <- data.frame(xcap1=prot$Yrot[,1], ycap1=prot$Yrot[,2], xcap2=prot$X[,1], ycap2=prot$X[,2])

#Add row names to df
plot$Sample <- row.names(plot)
#Merge according to Sample
Metadata2<-merge(Metadata2, plot, by="Sample")

#Change temperature and time to characters for plotting
Metadata2$Temperature<-as.character(Metadata2$Temperature)
Metadata2$Time<-as.character(Metadata2$Time)
#Change NAs in temperature to Direct
Metadata2$Temperature[is.na(Metadata2$Temperature)]<-"Direct"
#Change order for coloring
Metadata2$Temperature<-ordered(Metadata2$Temperature, levels=c("Direct", "-80", "-20", "5", "22"))

#Create plot name
pltName <- paste("Prot", i, sep="")
#create procrustes ggplot
ProcrustesList[[ pltName ]]<-ggplot(Metadata2) +
geom_point(aes(x=xcap1, y=ycap1, color=Temperature, shape = Time)) +
geom_point(aes(x=xcap2, y=ycap2, color=Temperature, shape = Time)) +
geom_segment(aes(x=xcap1, y=ycap1, xend=xcap2, yend=ycap2, color=Temperature), arrow=arrow(length=unit(0.2,"cm"))) +
scale_color_manual(values=c(Storagecol, Experimentcol)) + 
labs(colour="Temperature / °C", shape="Time / h", x = "Dimension 1", y = "Dimension 2", title = Title, subtitle = Subtitle) + 
theme_bw() + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12), legend.position="none")  
}

#Have the plots stored in lists
lay <- rbind(c(1,2,3,4),
             c(5,6,7,8))
pdf(paste("ProcrustesTAXvsAMR", ".pdf", sep=""), width=24, height=12)
grid.arrange(ProcrustesList$ProtP1spiked, ProcrustesList$ProtP2spiked, ProcrustesList$ProtS1spiked, ProcrustesList$ProtS2spiked, ProcrustesList$ProtP1unspiked, ProcrustesList$ProtP2unspiked, ProcrustesList$ProtS1unspiked, ProcrustesList$ProtS2unspiked, layout_matrix = lay)
dev.off()

#ProcrustesList$ProtP1spiked

lay <- rbind(c(1,5),
             c(2,6),
             c(3,7),
             c(4,8))
pdf(paste("ProcrustesTAXvsAMRTall", ".pdf", sep=""), width=8.5, height=11.75)
grid.arrange(ProcrustesList$ProtP1spiked, ProcrustesList$ProtP2spiked, ProcrustesList$ProtS1spiked, ProcrustesList$ProtS2spiked, ProcrustesList$ProtP1unspiked, ProcrustesList$ProtP2unspiked, ProcrustesList$ProtS1unspiked, ProcrustesList$ProtS2unspiked, layout_matrix = lay)
dev.off()

```

## Create MF5 and SF14
```{r}
#MF5 unspiked
lay <- rbind(c(1,2,3,4),
             c(1,2,3,4),
             c(5,5,6,6),
             c(5,5,6,6),
             c(5,5,6,6),
             c(7,7,8,8),
             c(7,7,8,8),
             c(7,7,8,8))
pdf(paste("MF5TotalAMRandProCrustes", ".pdf", sep=""), width=9, height=11.5)
grid.arrange(BoxplotListTotal$P1unspiked, BoxplotListTotal$P2unspiked, BoxplotListTotal$S1unspiked, BoxplotListTotal$S2unspiked, ProcrustesList$ProtP1unspiked, ProcrustesList$ProtP2unspiked, ProcrustesList$ProtS1unspiked, ProcrustesList$ProtS2unspiked, layout_matrix = lay)
dev.off()

#SF14 spiked
pdf(paste("SF14TotalAMRandProCrustes", ".pdf", sep=""), width=9, height=11.5)
grid.arrange(BoxplotListTotal$P1spiked, BoxplotListTotal$P2spiked, BoxplotListTotal$S1spiked, BoxplotListTotal$S2spiked, ProcrustesList$ProtP1spiked, ProcrustesList$ProtP2spiked, ProcrustesList$ProtS1spiked, ProcrustesList$ProtS2spiked, layout_matrix = lay)
dev.off()
```




## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```

