
---
title: "Dissimilarity assessment"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```


## Introduction
Full scale overview of dissimilarities (Supplementary Table 6) and relative to 0h in unspiked (Table 1)

### Packages 
```{r}
library(stringr) #Simple, consistent wrappers for common string operations. Like the str_replace function. help(package="stringr")
library(Rmisc) #The Rmisc library contains many functions useful for data analysis. Include summarySE help(package="Rmisc")
library(reshape2) #This function reshapes a data frame between 'wide' format with repeated measurements in separate columns of the same record and 'long' format with the repeated measurements in separate records help(package="reshape")
library(vegan) #The vegan package provides tools for descriptive community ecology. It has most basic functions of diversity analysis, community ordination and dissimilarity analysis. Most of its multivariate tools can be used for other data types as well. help(package="vegan")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr")
library(tidyr) #An evolution of 'reshape2'. It's designed specifically for data tidying (not general reshaping or aggregating) and works well with 'dplyr' data pipelines. help(package="tidyr")
```


## Define variables
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<- "All" #All, Allspiked, Allunspiked, PF, PFspiked, PFunspiked, SW, SWspiked, SWunspiked, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Vegdist dissimilarity method
Dist<-"bray" #manhattan, euclidean, canberra, bray, kulczynski, jaccard, gower, altGower, morisita, horn, mountford, raup, binomial, chao, cao or mahalanobis

#Decostand standardization method
Stand<-"hellinger" #total, max, freq, normalize, range, standardize, pa, chi.square, hellinger, log
```

##Analysis
### 1 Read in data
```{r}
Tax <- read.delim(file="Taxonomy20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```


### 2 Subsetting
```{r}
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting Metadata2
if (Subset=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (Subset=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (Subset=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (Subset=="PFspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked" | Sample_type == "Pig_feces_2_spiked")
} else if (Subset=="PFunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1" | Sample_type == "Pig_feces_2")
} else if (Subset=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (Subset=="SWspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked" | Sample_type == "Sewage_2_spiked")
} else if (Subset=="SWunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1" | Sample_type == "Sewage_2")
} else if (Subset=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (Subset=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (Subset=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (Subset=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (Subset=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (Subset=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (Subset=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (Subset=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (Subset=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (Subset=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (Subset=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (Subset=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (Subset=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
Tax2<-select(Tax, one_of(Metadata2$Sample))
```


### 3 Long formatted Bray Curtis dissimilarities
```{r}
#Multi dimensional scaling 
distmatrix <- vegdist(decostand(t(Tax2), method=Stand), method=Dist)

#Make distances into matrix
distmatrix<-as.matrix(distmatrix)

##Save table
#write.table(distmatrix, file="DistBrayHelAll.txt", quote=F, sep="\t")

#Contains all values except the ones equal to 0
meltPwBC<-subset(melt(distmatrix), value!=0)
rm(distmatrix)

#Have comparisons of both 1vs2 and 2vs1 ..., but are removed during the following filtration


```

### 4 All average dissimilarities unspiked (see below to run spiked)
```{r}
BCtable <- mutate(meltPwBC, from = str_replace(Var1,"[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_","") %>% str_replace("_[a-z]$",""))
BCtable <- mutate(BCtable, to = str_replace(Var2,"[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_","") %>% str_replace("_[a-z]$",""))
#Remove duplicate values, be careful could have same calculation that turned out the same, removes the three 0h comparisons both ways. 
#BCtable <- BCtable %>% distinct(value, .keep_all = TRUE)
#Calculate averages and dispersions
BCtable2 <- summarySE(BCtable, measurevar="value", groupvars=c("from", "to"))
#
BCtable2 <- select(BCtable2, one_of("from", "to", "value"))
#cast
BCtable2 <- spread(BCtable2, from, value)
#Vector of desired order
#cnam<-c("P1_0h", "P1_16h_80C", "P1_64h_80C", "P1_16h_20C", "P1_64h_20C", "P1_16h_5C", "P1_64h_5C", "P1_16h_22C", "P1_64h_22C", "P1_40h_2FTC_80C", "P1_64h_3FTC_80C", "P1_88h_4FTC_80C", "P1_40h_2FTC_20C", "P1_64h_3FTC_20C", "P1_88h_4FTC_20C", "P1_4m_80C", "P1_8m_80C", "P1_12m_80C", "P1_4m_20C", "P1_8m_20C", "P1_12m_20C", "KAHI_P1_0h", "KAHI_P1_64h_80C", "NFNS_P1_0h", "NFNS_P1_64h_80C", "P2_0h", "P2_16h_80C", "P2_64h_80C", "P2_16h_20C", "P2_64h_20C", "P2_16h_5C", "P2_64h_5C", "P2_16h_22C", "P2_64h_22C", "KAHI_P2_0h", "KAHI_P2_64h_80C", "NFNS_P2_0h", "NFNS_P2_64h_80C", "S1_0h", "S1_16h_80C", "S1_64h_80C", "S1_16h_20C", "S1_64h_20C", "S1_16h_5C", "S1_64h_5C", "S1_16h_22C", "S1_64h_22C",  "S1_40h_2FTC_80C", "S1_64h_3FTC_80C", "S1_88h_4FTC_80C", "S1_40h_2FTC_20C", "S1_64h_3FTC_20C", "S1_88h_4FTC_20C", "S1_4m_80C", "S1_8m_80C", "S1_12m_80C", "S1_4m_20C", "S1_8m_20C", "S1_12m_20C", "KAHI_S1_0h" , "KAHI_S1_64h_80C", "NFNS_S1_0h",  "NFNS_S1_64h_80C", "S2_0h", "S2_16h_80C", "S2_64h_80C", "S2_16h_20C", "S2_64h_20C", "S2_16h_5C", "S2_64h_5C", "S2_16h_22C", "S2_64h_22C", "KAHI_S2_0h", "KAHI_S2_64h_80C", "NFNS_S2_0h", "NFNS_S2_64h_80C") #All
cnam<-c("P1_0h", "P1_16h_80C", "P1_64h_80C", "P1_16h_20C", "P1_64h_20C", "P1_16h_5C", "P1_64h_5C", "P1_16h_22C", "P1_64h_22C", "P2_0h", "P2_16h_80C", "P2_64h_80C", "P2_16h_20C", "P2_64h_20C", "P2_16h_5C", "P2_64h_5C", "P2_16h_22C", "P2_64h_22C", "S1_0h", "S1_16h_80C", "S1_64h_80C", "S1_16h_20C", "S1_64h_20C", "S1_16h_5C", "S1_64h_5C", "S1_16h_22C", "S1_64h_22C", "S2_0h", "S2_16h_80C", "S2_64h_80C", "S2_16h_20C", "S2_64h_20C", "S2_16h_5C", "S2_64h_5C", "S2_16h_22C", "S2_64h_22C") #HX
#Order rows
BCtable2 <- BCtable2[match(cnam, BCtable2$to),]
#Make first column into rownames
BCtable2 <- data.frame(BCtable2[,-1], row.names=BCtable2[,1])
#Order columns
BCtable2<-BCtable2[cnam]

##Save table
write.table(BCtable2, file="DisBCAllavgUnspiked.txt", quote=F, sep="\t")
```

### 5 Dissimilarities relative to 0h, 0h is comparison within the replicates
```{r}
P10h<-BCtable2[c("P1_0h"),] %>% select(contains("P1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C")) 
P20h<-BCtable2[c("P2_0h"),] %>% select(contains("P2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S10h<-BCtable2[c("S1_0h"),] %>% select(contains("S1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S20h<-BCtable2[c("S2_0h"),] %>% select(contains("S2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))

Rel0h<-rbind(P10h, P20h, S10h, S20h)
write.table(Rel0h, file="DisBCAllavgUnspiked0h.txt", quote=F, sep="\t")
Rel0ht<-t(Rel0h)
write.table(Rel0ht, file="DisBCAllavgUnspiked0hTransposed.txt", quote=F, sep="\t")
```

### 6 Extract dispersion measurement
```{r}
#Calculate averages and dispersions
BCtableDisp <- summarySE(BCtable, measurevar="value", groupvars=c("from", "to"))
#
BCtableDisp <- select(BCtableDisp, one_of("from", "to", "se"))
#cast
BCtableDisp <- spread(BCtableDisp, from, se)
#Vector of desired order
#cnam<-c("P1_0h", "P1_16h_80C", "P1_64h_80C", "P1_16h_20C", "P1_64h_20C", "P1_16h_5C", "P1_64h_5C", "P1_16h_22C", "P1_64h_22C", "P1_40h_2FTC_80C", "P1_64h_3FTC_80C", "P1_88h_4FTC_80C", "P1_40h_2FTC_20C", "P1_64h_3FTC_20C", "P1_88h_4FTC_20C", "P1_4m_80C", "P1_8m_80C", "P1_12m_80C", "P1_4m_20C", "P1_8m_20C", "P1_12m_20C", "KAHI_P1_0h", "KAHI_P1_64h_80C", "NFNS_P1_0h", "NFNS_P1_64h_80C", "P2_0h", "P2_16h_80C", "P2_64h_80C", "P2_16h_20C", "P2_64h_20C", "P2_16h_5C", "P2_64h_5C", "P2_16h_22C", "P2_64h_22C", "KAHI_P2_0h", "KAHI_P2_64h_80C", "NFNS_P2_0h", "NFNS_P2_64h_80C", "S1_0h", "S1_16h_80C", "S1_64h_80C", "S1_16h_20C", "S1_64h_20C", "S1_16h_5C", "S1_64h_5C", "S1_16h_22C", "S1_64h_22C",  "S1_40h_2FTC_80C", "S1_64h_3FTC_80C", "S1_88h_4FTC_80C", "S1_40h_2FTC_20C", "S1_64h_3FTC_20C", "S1_88h_4FTC_20C", "S1_4m_80C", "S1_8m_80C", "S1_12m_80C", "S1_4m_20C", "S1_8m_20C", "S1_12m_20C", "KAHI_S1_0h" , "KAHI_S1_64h_80C", "NFNS_S1_0h",  "NFNS_S1_64h_80C", "S2_0h", "S2_16h_80C", "S2_64h_80C", "S2_16h_20C", "S2_64h_20C", "S2_16h_5C", "S2_64h_5C", "S2_16h_22C", "S2_64h_22C", "KAHI_S2_0h", "KAHI_S2_64h_80C", "NFNS_S2_0h", "NFNS_S2_64h_80C") #All
cnam<-c("P1_0h", "P1_16h_80C", "P1_64h_80C", "P1_16h_20C", "P1_64h_20C", "P1_16h_5C", "P1_64h_5C", "P1_16h_22C", "P1_64h_22C", "P2_0h", "P2_16h_80C", "P2_64h_80C", "P2_16h_20C", "P2_64h_20C", "P2_16h_5C", "P2_64h_5C", "P2_16h_22C", "P2_64h_22C", "S1_0h", "S1_16h_80C", "S1_64h_80C", "S1_16h_20C", "S1_64h_20C", "S1_16h_5C", "S1_64h_5C", "S1_16h_22C", "S1_64h_22C", "S2_0h", "S2_16h_80C", "S2_64h_80C", "S2_16h_20C", "S2_64h_20C", "S2_16h_5C", "S2_64h_5C", "S2_16h_22C", "S2_64h_22C") #HX
#Order rows
BCtableDisp <- BCtableDisp[match(cnam, BCtableDisp$to),]
#Make first column into rownames
BCtableDisp <- data.frame(BCtableDisp[,-1], row.names=BCtableDisp[,1])
#Order columns
BCtableDisp<-BCtableDisp[cnam]

##Save table
write.table(BCtableDisp, file="DisBCAllavgUnspikedSE.txt", quote=F, sep="\t")
```

### 7 Dispersion relative to 0h, 0h is comparison within the replicates
```{r}
P10h<-BCtableDisp[c("P1_0h"),] %>% select(contains("P1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C")) 
P20h<-BCtableDisp[c("P2_0h"),] %>% select(contains("P2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S10h<-BCtableDisp[c("S1_0h"),] %>% select(contains("S1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S20h<-BCtableDisp[c("S2_0h"),] %>% select(contains("S2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))

Rel0h<-rbind(P10h, P20h, S10h, S20h)
write.table(Rel0h, file="DisBCAllavgUnspiked0hSE.txt", quote=F, sep="\t")
Rel0ht<-t(Rel0h)
write.table(Rel0ht, file="DisBCAllavgUnspiked0hTransposedSE.txt", quote=F, sep="\t")
```

### 8 All average dissimilarities spiked
```{r}
BCtable <- mutate(meltPwBC, from = str_replace(Var1,"[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_","") %>% str_replace("_[a-z]$",""))
BCtable <- mutate(BCtable, to = str_replace(Var2,"[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_","") %>% str_replace("_[a-z]$",""))
#Remove duplicate values, be careful could have same calculation that turned out the same, removes the three 0h comparisons both ways. 
#BCtable <- BCtable %>% distinct(value, .keep_all = TRUE)
#Calculate averages and dispersions
BCtable2 <- summarySE(BCtable, measurevar="value", groupvars=c("from", "to"))
#
BCtable2 <- select(BCtable2, one_of("from", "to", "value"))
#cast
BCtable2 <- spread(BCtable2, from, value)
#Vector of desired order
#cnam<-c("P1MC_0h", "P1MC_16h_80C", "P1MC_64h_80C", "P1MC_16h_20C", "P1MC_64h_20C", "P1MC_16h_5C", "P1MC_64h_5C", "P1MC_16h_22C", "P1MC_64h_22C", "P1MC_40h_2FTC_80C", "P1MC_64h_3FTC_80C", "P1MC_88h_4FTC_80C", "P1MC_40h_2FTC_20C", "P1MC_64h_3FTC_20C", "P1MC_88h_4FTC_20C", "P1MC_4m_80C", "P1MC_8m_80C", "P1MC_12m_80C", "P1MC_4m_20C", "P1MC_8m_20C", "P1MC_12m_20C", "P2MC_0h", "P2MC_16h_80C", "P2MC_64h_80C", "P2MC_16h_20C", "P2MC_64h_20C", "P2MC_16h_5C", "P2MC_64h_5C", "P2MC_16h_22C", "P2MC_64h_22C", "S1MC_0h", "S1MC_16h_80C", "S1MC_64h_80C", "S1MC_16h_20C", "S1MC_64h_20C", "S1MC_16h_5C", "S1MC_64h_5C", "S1MC_16h_22C", "S1MC_64h_22C",  "S1MC_40h_2FTC_80C", "S1MC_64h_3FTC_80C", "S1MC_88h_4FTC_80C", "S1MC_40h_2FTC_20C", "S1MC_64h_3FTC_20C", "S1MC_88h_4FTC_20C", "S1MC_4m_80C", "S1MC_8m_80C", "S1MC_12m_80C", "S1MC_4m_20C", "S1MC_8m_20C", "S1MC_12m_20C", "S2MC_0h", "S2MC_16h_80C", "S2MC_64h_80C", "S2MC_16h_20C", "S2MC_64h_20C", "S2MC_16h_5C", "S2MC_64h_5C", "S2MC_16h_22C", "S2MC_64h_22C") #All
cnam<-c("P1MC_0h", "P1MC_16h_80C", "P1MC_64h_80C", "P1MC_16h_20C", "P1MC_64h_20C", "P1MC_16h_5C", "P1MC_64h_5C", "P1MC_16h_22C", "P1MC_64h_22C", "P2MC_0h", "P2MC_16h_80C", "P2MC_64h_80C", "P2MC_16h_20C", "P2MC_64h_20C", "P2MC_16h_5C", "P2MC_64h_5C", "P2MC_16h_22C", "P2MC_64h_22C", "S1MC_0h", "S1MC_16h_80C", "S1MC_64h_80C", "S1MC_16h_20C", "S1MC_64h_20C", "S1MC_16h_5C", "S1MC_64h_5C", "S1MC_16h_22C", "S1MC_64h_22C", "S2MC_0h", "S2MC_16h_80C", "S2MC_64h_80C", "S2MC_16h_20C", "S2MC_64h_20C", "S2MC_16h_5C", "S2MC_64h_5C", "S2MC_16h_22C", "S2MC_64h_22C") #HX
#Order rows
BCtable2 <- BCtable2[match(cnam, BCtable2$to),]
#Make first column into rownames
BCtable2 <- data.frame(BCtable2[,-1], row.names=BCtable2[,1])
#Order columns
BCtable2<-BCtable2[cnam]

#Save table
write.table(BCtable2, file="DisBCAllavgSpiked.txt", quote=F, sep="\t")
```

### 9 Dissimilarities relative to 0h, 0h is comparison within the replicates
```{r}
P10h<-BCtable2[c("P1_0h"),] %>% select(contains("P1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C")) 
P20h<-BCtable2[c("P2_0h"),] %>% select(contains("P2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S10h<-BCtable2[c("S1_0h"),] %>% select(contains("S1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S20h<-BCtable2[c("S2_0h"),] %>% select(contains("S2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))

Rel0h<-rbind(P10h, P20h, S10h, S20h)
write.table(Rel0h, file="DisBCAllavgSpiked0h.txt", quote=F, sep="\t")
Rel0ht<-t(Rel0h)
write.table(Rel0ht, file="DisBCAllavgSpiked0hTransposed.txt", quote=F, sep="\t")
```

### 10 Extract dispersion measurement
```{r}
#Calculate averages and dispersions
BCtableDisp <- summarySE(BCtable, measurevar="value", groupvars=c("from", "to"))
#
BCtableDisp <- select(BCtableDisp, one_of("from", "to", "se"))
#cast
BCtableDisp <- spread(BCtableDisp, from, se)
#Vector of desired order
#cnam<-c("P1_0h", "P1_16h_80C", "P1_64h_80C", "P1_16h_20C", "P1_64h_20C", "P1_16h_5C", "P1_64h_5C", "P1_16h_22C", "P1_64h_22C", "P1_40h_2FTC_80C", "P1_64h_3FTC_80C", "P1_88h_4FTC_80C", "P1_40h_2FTC_20C", "P1_64h_3FTC_20C", "P1_88h_4FTC_20C", "P1_4m_80C", "P1_8m_80C", "P1_12m_80C", "P1_4m_20C", "P1_8m_20C", "P1_12m_20C", "KAHI_P1_0h", "KAHI_P1_64h_80C", "NFNS_P1_0h", "NFNS_P1_64h_80C", "P2_0h", "P2_16h_80C", "P2_64h_80C", "P2_16h_20C", "P2_64h_20C", "P2_16h_5C", "P2_64h_5C", "P2_16h_22C", "P2_64h_22C", "KAHI_P2_0h", "KAHI_P2_64h_80C", "NFNS_P2_0h", "NFNS_P2_64h_80C", "S1_0h", "S1_16h_80C", "S1_64h_80C", "S1_16h_20C", "S1_64h_20C", "S1_16h_5C", "S1_64h_5C", "S1_16h_22C", "S1_64h_22C",  "S1_40h_2FTC_80C", "S1_64h_3FTC_80C", "S1_88h_4FTC_80C", "S1_40h_2FTC_20C", "S1_64h_3FTC_20C", "S1_88h_4FTC_20C", "S1_4m_80C", "S1_8m_80C", "S1_12m_80C", "S1_4m_20C", "S1_8m_20C", "S1_12m_20C", "KAHI_S1_0h" , "KAHI_S1_64h_80C", "NFNS_S1_0h",  "NFNS_S1_64h_80C", "S2_0h", "S2_16h_80C", "S2_64h_80C", "S2_16h_20C", "S2_64h_20C", "S2_16h_5C", "S2_64h_5C", "S2_16h_22C", "S2_64h_22C", "KAHI_S2_0h", "KAHI_S2_64h_80C", "NFNS_S2_0h", "NFNS_S2_64h_80C") #All
cnam<-c("P1_0h", "P1_16h_80C", "P1_64h_80C", "P1_16h_20C", "P1_64h_20C", "P1_16h_5C", "P1_64h_5C", "P1_16h_22C", "P1_64h_22C", "P2_0h", "P2_16h_80C", "P2_64h_80C", "P2_16h_20C", "P2_64h_20C", "P2_16h_5C", "P2_64h_5C", "P2_16h_22C", "P2_64h_22C", "S1_0h", "S1_16h_80C", "S1_64h_80C", "S1_16h_20C", "S1_64h_20C", "S1_16h_5C", "S1_64h_5C", "S1_16h_22C", "S1_64h_22C", "S2_0h", "S2_16h_80C", "S2_64h_80C", "S2_16h_20C", "S2_64h_20C", "S2_16h_5C", "S2_64h_5C", "S2_16h_22C", "S2_64h_22C") #HX
#Order rows
BCtableDisp <- BCtableDisp[match(cnam, BCtableDisp$to),]
#Make first column into rownames
BCtableDisp <- data.frame(BCtableDisp[,-1], row.names=BCtableDisp[,1])
#Order columns
BCtableDisp<-BCtableDisp[cnam]

##Save table
write.table(BCtableDisp, file="DisBCAllavgSpikedSE.txt", quote=F, sep="\t")
```

### 11 Dispersion relative to 0h, 0h is comparison within the replicates
```{r}
P10h<-BCtableDisp[c("P1_0h"),] %>% select(contains("P1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C")) 
P20h<-BCtableDisp[c("P2_0h"),] %>% select(contains("P2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S10h<-BCtableDisp[c("S1_0h"),] %>% select(contains("S1")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))
S20h<-BCtableDisp[c("S2_0h"),] %>% select(contains("S2")) %>% setNames(., c("0h", "16h_80C", "64h_80C", "16h_20C", "64h_20C", "16h_5C", "64h_5C", "16h_22C", "64h_22C"))

Rel0h<-rbind(P10h, P20h, S10h, S20h)
write.table(Rel0h, file="DisBCAllavgSpiked0hSE.txt", quote=F, sep="\t")
Rel0ht<-t(Rel0h)
write.table(Rel0ht, file="DisBCAllavgSpiked0hTransposedSE.txt", quote=F, sep="\t")
```


## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```

