---
title: "Create Stacked bar charts mock community organisms"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
Creating stacked bar charts from mock community organisms.  
The data were created from SF10andST7MCorgsStackedBar1.Rmd and data from microscopy counts, CFU counts, and DNA isolation concentrations. See ST7 for all data structured to run this script. 

### Packages
```{r}
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2")
```


## Analysis
### 1 Create stacked bar charts 
```{r}
Spike <- read.delim(file="ST7StackedMCorgs.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")

#Change order of boxplots on x-axis
Spike$Type<-as.factor(Spike$Type)
levels(Spike$Type)
Spike$Type<-ordered(Spike$Type, levels=c("Mic", "CFU", "Con","0h", "16h-80C", "64h-80C", "16h-20C", "64h-20C", "16h5C", "64h5C", "16h22C", "64h22C")) #Have to change according to Type #17_9 "Mic", "CFU", "Con", "GSRelMC", "DESeqGSRel0h", "DESeqGSRel16h80C", "DESeqGSRel16h20C", "DESeqGSRel16h5C", "DESeqGSRel16h22C", "DESeqGSRel64h80C", "DESeqGSRel64h20C", "DESeqGSRel64h5C", "DESeqGSRel64h22C" #17_1 "Mic", "CFU", "Con", "RelMC", "RelS", "GSRelMC", "GSRelS", "GSHelMC", "GSHelS", "GSdWiMC", "GSdWiS", "GSlogMC", "GSlogS", "GSSchisMC", "GSchisS"
levels(Spike$Type)
#Change order of Organisms
Spike$Organism<-as.factor(Spike$Organism)
levels(Spike$Organism)
Spike$Organism<-ordered(Spike$Organism, levels=c("Propionibacterium", "Staphylococcus", "Bacteroides","Fusobacterium", "Escherichia", "Salmonella", "Saccharomyces", "Cryptosporidium"))
levels(Spike$Organism)

ggplot(Spike, aes(x = Type, y = Value_percent, fill=Organism)) + facet_wrap(~Facet) + geom_bar(stat = "identity") +
scale_fill_brewer(palette="Set1") + xlab("Type") + ylab("Percentage (%)")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))

pdf("SF10StackedBarMCOrg.pdf", height=8, width=12)
ggplot(Spike, aes(x = Type, y = Value_percent, fill=Organism)) + 
  facet_wrap(~Facet) + geom_bar(stat = "identity") +
  scale_fill_brewer(palette="Set1") + 
  xlab("Type") + 
  ylab("Percentage (%)")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") + 
  guides(fill = guide_legend(nrow = 1))
dev.off()

png("SF10StackedBarMCOrg.png", height=800, width=1200)
ggplot(Spike, aes(x = Type, y = Value_percent, fill=Organism)) + 
  facet_wrap(~Facet) + geom_bar(stat = "identity") +
  scale_fill_brewer(palette="Set1") + 
  xlab("Type") + 
  ylab("Percentage (%)")+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") + 
  guides(fill = guide_legend(nrow = 1))
dev.off()
```

## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```




