---
title: "Rarecurves AMR"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

## Introduction
Creating rarefaction curves to evaluate if the genera detected are reaching a plateu indicating that samples have been sequenced at a level where only rare genera remains undetected.  
Same as "SF1TaxRarecurves.Rmd" except changing the raw count table to the raw reads obtained from the AMR genes and the step=10 in the rarecurve function. Name y-axis accordingly. 

### Packages 
```{r}
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2")  
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr"))
library(rtk) #Rarefy data, calculate diversity and plot the results  
```

## Define variables used in the analysis
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<- "All" #All, Allspiked, Allunspiked, PF, SW, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. 
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both
```



## Analysis
### 1 Read in data
```{r}
TaxRaw <- read.delim(file="AMRRaw20180919.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```

### 2 Subsetting 
```{r}
#Removing negative and positive controls
Metadata<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting metadata
if (Subset=="Allspiked") {
  Metadata<-filter(Metadata, SpikedUnspiked == "Spiked")
} else if (Subset=="Allunspiked") {
  Metadata<-filter(Metadata, SpikedUnspiked == "Unspiked")
} else if (Subset=="PF") {
  Metadata<-filter(Metadata, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (Subset=="SW") {
  Metadata<-filter(Metadata, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (Subset=="P1") {
  Metadata<-filter(Metadata, Experiment == "Pig_feces_1")
} else if (Subset=="P1spiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_1_spiked")
} else if (Subset=="P1unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_1")
} else if (Subset=="P2") {
  Metadata<-filter(Metadata, Experiment == "Pig_feces_2")
} else if (Subset=="P2spiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_2_spiked")
} else if (Subset=="P2unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_2")
} else if (Subset=="S1") {
  Metadata<-filter(Metadata, Experiment == "Sewage_1")
} else if (Subset=="S1spiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_1_spiked")
} else if (Subset=="S1unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_1")
} else if (Subset=="S2") {
  Metadata<-filter(Metadata, Experiment == "Sewage_2")
} else if (Subset=="S2spiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_2_spiked")
} else if (Subset=="S2unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_2")
} else if (Subset=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubExp=="HX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata<-filter(Metadata, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata<-filter(Metadata, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata2<-filter(Metadata, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata2$Matching_samples)
  Metadata<-filter(Metadata, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata2)
} else if (SubExp=="HXFTX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubFre=="Frozen") {
  Metadata<-filter(Metadata, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata<-filter(Metadata, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

TaxRaw<-select(TaxRaw, one_of(c(Metadata$Sample)))
```


### 3 rarefaction curves
```{r}
rare<-data.frame(t(TaxRaw))
rare<-add_rownames(rare, "Sample")
rare$colors<-ifelse(grepl("_P1", rare$Sample), "#30241E", ifelse(grepl("_P2", rare$Sample), "#B59B80", ifelse(grepl("_S1", rare$Sample), "#33a02c", ifelse(grepl("_S2", rare$Sample), "#B2DF8A", "pink"))))

rarecurve(t(TaxRaw), step=10, xlab="Sample size", ylab="Genera", col=rare$colors, label=FALSE)
#Does not work  , legend(10000000, 4000, c("P1", "P2", "S1", "S2"), lty=c(1,1), lwd=c(2.5,2.5), col=rare$colors)

pdf(paste("SF12AMR", "rarecurveAll.pdf", sep=""), height=6, width=8)
rarecurve(t(TaxRaw), step=10, xlab="Mapped reads", ylab="Genes", col=rare$colors, label=FALSE)
dev.off()

png(paste("SF12AMR", "rarecurveAll.png", sep=""), height=300, width=400)
rarecurve(t(TaxRaw), step=10, xlab="Mapped reads", ylab="Genes", col=rare$colors, label=FALSE)
dev.off()


#Can run to colour according to storage conditions when subsetting to P1, P2, S1, S2.
#P1
TaxRawP1<-select(TaxRaw, matches("_P1"))
rare<-data.frame(t(TaxRawP1))
rare<-add_rownames(rare, "Sample")
#________________
#Different colour stategies
#Temp
rare$colors<-ifelse(grepl("0h", rare$Sample), "#663300", ifelse(grepl("5C", rare$Sample), "#FF9900", ifelse(grepl("22C", rare$Sample), "#006600", ifelse(grepl("0C", rare$Sample), "#33FF33", "pink"))))
#Spiked or unspiked
rare$colors<-ifelse(grepl("MC", rare$Sample), "blue", "red")
#________________
#pdf(paste("Genus", "rarecurveP1.pdf", sep=""), height=6, width=12)
rarecurve(t(TaxRawP1), step=10, xlab="Sample size", ylab="Genus", col=rare$colors, label=FALSE)
#dev.off()

#Can run to colour according to storage conditions when subsetting to P1, P2, S1, S2.
#P2
TaxRawP2<-select(TaxRaw, matches("_P2"))
rare<-data.frame(t(TaxRawP2))
rare<-add_rownames(rare, "Sample")
#________________
#Different colour stategies
#Temp
rare$colors<-ifelse(grepl("0h", rare$Sample), "#663300", ifelse(grepl("5C", rare$Sample), "#FF9900", ifelse(grepl("22C", rare$Sample), "#006600", ifelse(grepl("0C", rare$Sample), "#33FF33", "pink"))))
#Spiked or unspiked
rare$colors<-ifelse(grepl("MC", rare$Sample), "blue", "red")
#________________
#pdf(paste("Genus", "rarecurveP2.pdf", sep=""), height=6, width=12)
rarecurve(t(TaxRawP2), step=10, xlab="Sample size", ylab="Genus", col=rare$colors, label=FALSE)
#dev.off()

#Can run to colour according to storage conditions when subsetting to P1, P2, S1, S2.
#S1
TaxRawS1<-select(TaxRaw, matches("_S1"))
rare<-data.frame(t(TaxRawS1))
rare<-add_rownames(rare, "Sample")
#________________
#Different colour stategies
#Temp
rare$colors<-ifelse(grepl("0h", rare$Sample), "#663300", ifelse(grepl("5C", rare$Sample), "#FF9900", ifelse(grepl("22C", rare$Sample), "#006600", ifelse(grepl("0C", rare$Sample), "#33FF33", "pink"))))
#Spiked or unspiked
rare$colors<-ifelse(grepl("MC", rare$Sample), "blue", "red")
#________________
#pdf(paste("Genus", "rarecurveS1.pdf", sep=""), height=6, width=12)
rarecurve(t(TaxRawS1), step=10, xlab="Sample size", ylab="Genus", col=rare$colors, label=FALSE)
#dev.off()

#Can run to colour according to storage conditions when subsetting to P1, P2, S1, S2.
#S2
TaxRawS2<-select(TaxRaw, matches("_S2"))
rare<-data.frame(t(TaxRawS2))
rare<-add_rownames(rare, "Sample")
#________________
#Different colour stategies
#Temp
rare$colors<-ifelse(grepl("0h", rare$Sample), "#663300", ifelse(grepl("5C", rare$Sample), "#FF9900", ifelse(grepl("22C", rare$Sample), "#006600", ifelse(grepl("0C", rare$Sample), "#33FF33", "pink"))))
#Spiked or unspiked
rare$colors<-ifelse(grepl("MC", rare$Sample), "blue", "red")
#________________
#pdf(paste("Genus", "rarecurveS2.pdf", sep=""), height=6, width=12)
rarecurve(t(TaxRawS2), step=10, xlab="Sample size", ylab="Genus", col=rare$colors, label=FALSE)
#dev.off()
```

## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```
