
---
title: "Alpha-diversity"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
Calculating alpha diversity measures for pig feces and sewage.  
Indices include: Shannon. Simpson, inverse Simpson, Fishers alpha, richness, rarefy_min_count, chao1, chao1SE, Shannon with rarefy, Simpson with rarefy, Inverse Simpson with rarefy, and Pielou's evenness.   

### Packages
```{r}
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2")  
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr")
library(gridExtra)
library(reshape2)
library(tidyr)
```

## Define variables 
Filter rows is in this case due to subsetting getting rows only containing zeroes 
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<-"All" #All, Allspiked, Allunspiked, PF, SW, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All #Need to modify to run with other experiments than HX

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Filtering of the count table.
Filter_Rows<-0 #Is based on row sums so can take integer values. Pick zero to remove not present genera relevant due to the subsetting 
```


## Analysis Pig feces
### 1 Read in data
```{r}
TaxRaw <- read.delim(file="TaxonomyRaw20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```


### 2 Subsetting 
```{r}
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting metadata
if (Subset=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (Subset=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (Subset=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (Subset=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (Subset=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (Subset=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (Subset=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (Subset=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (Subset=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (Subset=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (Subset=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (Subset=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (Subset=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (Subset=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (Subset=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (Subset=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (Subset=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
TaxRaw2<-select(TaxRaw, one_of(Metadata2$Sample))

#Change Time to be a factor
#Metadata2$Time<-as.factor(Metadata2$Time)
```


### 3 Filtering
```{r}
##Remove spiked in orgs at genus level
#SpikedGenus<-c("Escherichia", "Bacteroides", "Salmonella", "Fusobacterium", "Propionibacterium", "Staphylococcus", "Cryptosporidium", "Saccharomyces")
#TaxRaw2$Orgs<-rownames(TaxRaw2)
#TaxRaw2<-TaxRaw2[!TaxRaw2$Orgs %in% SpikedGenus, ]
#TaxRaw2<-TaxRaw2[, 1:length(TaxRaw2)-1]

#Filtering of the Counttable depending on rowSums. 
TaxRaw2 <- TaxRaw2[rowSums(TaxRaw2)>Filter_Rows*length(TaxRaw2),]
```



### 4 Calculate alpha diversity and add to metadata
```{r}
#Use vegan to calculate various diversity and richness indexes for each sample
diversityCalc <- data.frame(Shannon=diversity(t(TaxRaw2), index="shannon"), Simpson=diversity(t(TaxRaw2), index="simpson"), invSimpson=diversity(t(TaxRaw2), index="invsimpson"), fisher=fisher.alpha(t(TaxRaw2)), richness=specnumber(t(TaxRaw2)), rarefy_min_count=rarefy(t(TaxRaw2), sample=min(rowSums(t(TaxRaw2)))), chao1=estimateR(t(TaxRaw2))["S.chao1",], chao1SE=estimateR(t(TaxRaw2))["se.chao1",], ShannonRar=diversity(rrarefy(data.frame(t(TaxRaw2)), min(rowSums(t(TaxRaw2)))), index="shannon"), SimpsonRar=diversity(rrarefy(data.frame(t(TaxRaw2)), min(rowSums(t(TaxRaw2)))), index="simpson"), invSimpsonRar=diversity(rrarefy(data.frame(t(TaxRaw2)), min(rowSums(t(TaxRaw2)))), index="invsimpson"), Pielou=diversity(t(TaxRaw2))/log(specnumber(t(TaxRaw2))))

diversityCalc<-add_rownames(diversityCalc, "Sample")

Metadata2<-merge(Metadata2, diversityCalc, by="Sample")
```


### 5 Make boxplots
```{r}
##Select organisms
AlphaDiv<-c("fisher", "Shannon", "Simpson", "invSimpson", "richness", "chao1", "ShannonRar", "SimpsonRar", "invSimpsonRar", "Pielou")
Metadata2$Replicate_Boxplot<-as.factor(Metadata2$Replicate_Boxplot)
#Change order on x-axis
levels(Metadata2$Replicate_Boxplot)
Metadata2$Replicate_Boxplot<-ordered(Metadata2$Replicate_Boxplot, levels=c("0h", "16h-80C", "64h-80C", "16h-20C", "64h-20C", "16h5C", "64h5C", "16h22C", "64h22C"))
Metadata2$SpikedUnspiked<-ordered(Metadata2$SpikedUnspiked, levels=c("Unspiked", "Spiked"))

# Create a list to hold the plot objects.
AlphaList <- list()

#Plot
for (i in AlphaDiv) {
#Create plot name
pltName <- paste( 'PFAlpha', i, sep = '' )
#create PCoA
AlphaList[[ pltName ]]<- 
ggplot(Metadata2, aes_string(x="Replicate_Boxplot", y=i, group="Replicate_Boxplot")) +
    geom_boxplot(aes(fill=Replicate_Boxplot)) + 
    facet_grid(SpikedUnspiked ~ Sample_type_new, scales="fixed") + #Scales should not be "free"
    ggtitle(paste("Genus", Subset, SubExp, i, sep=" ")) + 
    xlab("Storage conditions") + 
    ylab("Index value") + 
    scale_fill_manual(values=c("#999999", "#0571b0", "#0571b0", "#92c5de", "#92c5de", "#f4a582", "#f4a582", "#ca0020", "#ca0020")) + 
    theme_bw() + 
    theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1))
}

mean(Metadata2$chao1)
sd(Metadata2$chao1)

fropilmean<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer") 
mean(fropilmean$Pielou)
sd(fropilmean$Pielou)

Unfropilmean<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
mean(Unfropilmean$Pielou)
sd(Unfropilmean$Pielou)
```



### 6 Sewage analysis
Same analysis as pig feces. Using the same data
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<-"SW" #All, Allspiked, Allunspiked, PF, SW, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting metadata
if (Subset=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (Subset=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (Subset=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (Subset=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (Subset=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (Subset=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (Subset=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (Subset=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (Subset=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (Subset=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (Subset=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (Subset=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (Subset=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (Subset=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (Subset=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (Subset=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (Subset=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
TaxRaw2<-select(TaxRaw, one_of(Metadata2$Sample))

#Filtering of the Counttable depending on rowSums. 
TaxRaw2 <- TaxRaw2[rowSums(TaxRaw2)>Filter_Rows*length(TaxRaw2),]


#Calculate alpha diversity and add to metadata
#Use vegan to calculate various diversity and richness indexes for each sample
diversityCalc <- data.frame(Shannon=diversity(t(TaxRaw2), index="shannon"), Simpson=diversity(t(TaxRaw2), index="simpson"), invSimpson=diversity(t(TaxRaw2), index="invsimpson"), fisher=fisher.alpha(t(TaxRaw2)), richness=specnumber(t(TaxRaw2)), rarefy_min_count=rarefy(t(TaxRaw2), sample=min(rowSums(t(TaxRaw2)))), chao1=estimateR(t(TaxRaw2))["S.chao1",], chao1SE=estimateR(t(TaxRaw2))["se.chao1",], ShannonRar=diversity(rrarefy(data.frame(t(TaxRaw2)), min(rowSums(t(TaxRaw2)))), index="shannon"), SimpsonRar=diversity(rrarefy(data.frame(t(TaxRaw2)), min(rowSums(t(TaxRaw2)))), index="simpson"), invSimpsonRar=diversity(rrarefy(data.frame(t(TaxRaw2)), min(rowSums(t(TaxRaw2)))), index="invsimpson"), Pielou=diversity(t(TaxRaw2))/log(specnumber(t(TaxRaw2))))

diversityCalc<-add_rownames(diversityCalc, "Sample")

Metadata2<-merge(Metadata2, diversityCalc, by="Sample")

#Make boxplots
##Select organisms
AlphaDiv<-c("fisher", "Shannon", "Simpson", "invSimpson", "richness", "chao1", "ShannonRar", "SimpsonRar", "invSimpsonRar", "Pielou")
Metadata2$Replicate_Boxplot<-as.factor(Metadata2$Replicate_Boxplot)
#Change order on x-axis
levels(Metadata2$Replicate_Boxplot)
Metadata2$Replicate_Boxplot<-ordered(Metadata2$Replicate_Boxplot, levels=c("0h", "16h-80C", "64h-80C", "16h-20C", "64h-20C", "16h5C", "64h5C", "16h22C", "64h22C"))
Metadata2$SpikedUnspiked<-ordered(Metadata2$SpikedUnspiked, levels=c("Unspiked", "Spiked"))

#Plot
for (i in AlphaDiv) {
#Create plot name
pltName <- paste( 'SWAlpha', i, sep = '' )
#create PCoA
AlphaList[[ pltName ]]<- 
ggplot(Metadata2, aes_string(x="Replicate_Boxplot", y=i, group="Replicate_Boxplot")) +
    geom_boxplot(aes(fill=Replicate_Boxplot)) + 
    facet_grid(SpikedUnspiked ~ Sample_type_new, scales="fixed") + #Scales should not be "free"
    ggtitle(paste("Genus", Subset, SubExp, i, sep=" ")) + 
    xlab("Storage conditions") + 
    ylab("Index value") + 
    scale_fill_manual(values=c("#999999", "#0571b0", "#0571b0", "#92c5de", "#92c5de", "#f4a582", "#f4a582", "#ca0020", "#ca0020")) + 
    theme_bw() + 
    theme(legend.position="none", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1))
}

mean(Metadata2$chao1)
sd(Metadata2$chao1)

```

### 7 Create figure with selected indices 
```{r}
#Have the plots stored in lists
lay <- rbind(c(1,2),
             c(3,4),
             c(5,6))
pdf(paste("SF2Alpha", "Genus", ".pdf", sep=""), width=8, height=12)
grid.arrange(AlphaList$PFAlphachao1, AlphaList$SWAlphachao1, AlphaList$PFAlphaPielou, AlphaList$SWAlphaPielou, AlphaList$PFAlphaSimpson, AlphaList$SWAlphaSimpson, layout_matrix = lay)
dev.off()

png(paste("SF2Alpha", "Genus", ".png", sep=""), width=600, height=900)
grid.arrange(AlphaList$PFAlphachao1, AlphaList$SWAlphachao1, AlphaList$PFAlphaPielou, AlphaList$SWAlphaPielou, AlphaList$PFAlphaSimpson, AlphaList$SWAlphaSimpson, layout_matrix = lay)
dev.off()
```


#_________________________________________________________________
###### Session information
```{r session_info}
sessionInfo()

```

######This document was processed on: 
```{r}
Sys.Date()
```




