---
title: "Statistical assessment of the dissimilarities"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
The purpose of the following code is to perform a statistical assessment of the pairwise dissimilarities. The purpose is to supplement the 'adonis' analysis that revealed some groups had a significant difference in group homogeneities.  

Parametric methods are not appropriate therefore a nonparametric test were implemented. 

### Packages
```{r}
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2")  
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
help(package="dplyr")
library(stringr) #Simple, consistent wrappers for common string operations. Like the str_replace function. help(package="stringr")
library(reshape2) # Flexibly restructure and aggregate data using just two functions: melt and 'dcast' (or 'acast'). help(package="reshape2")
library(car) # Companion to applied regression includes Levene's test. help(package="car")
library(FSA) #A variety of simple fish stock assessment methods. Includes the Dunne non-parametric follow-up test help(package="FSA")
```


## Define variables
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<-"Allspiked" #All, Allspiked, Allunspiked, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Vegdist dissimilarity method
Dist<-"bray" #manhattan, euclidean, canberra, bray, kulczynski, jaccard, gower, altGower, morisita, horn, mountford, raup, binomial, chao, cao or mahalanobis

#Decostand standardization method
Stand<-"hellinger" #total, max, freq, normalize, range, standardize, pa, chi.square, hellinger, log

#If the ANOVA is significant which post test to use
PostTestParametric<-"ttest" #ttest, TukeyHSD

#If PostTestParametric<-"ttest", which p-value correction to use
padj<-"bonferroni" #holm, hochberg, hommel, bonferroni, BH, BY, fdr, none

#If the Kruskal is significant which post test to use
PostTestNonParametric<-"Dunne" #Dunne, MannWhitney

#If PostTestNonParametric<-"Dunne", which p-value correction to use
Dunnepadj<-"bonferroni" #holm, hochberg, hommel, bonferroni, BH, BY, fdr, none 

#If PostTestNonParametric<-"MannWhitney", which p-value correction to use
MWpadj<-"bonferroni" #holm, hochberg, hommel, bonferroni, BH, BY, fdr, none
```

## Analysis
### 1 Read in data
```{r}
Tax <- read.delim(file="Taxonomy20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```

### 2 Subsetting
```{r}
#Removing negative and positive controls
Metadata<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting metadata
if (Subset=="Allspiked") {
  Metadata<-filter(Metadata, SpikedUnspiked == "Spiked")
} else if (Subset=="Allunspiked") {
  Metadata<-filter(Metadata, SpikedUnspiked == "Unspiked")
} else if (Subset=="P1") {
  Metadata<-filter(Metadata, Experiment == "Pig_feces_1")
} else if (Subset=="P1spiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_1_spiked")
} else if (Subset=="P1unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_1")
} else if (Subset=="P2") {
  Metadata<-filter(Metadata, Experiment == "Pig_feces_2")
} else if (Subset=="P2spiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_2_spiked")
} else if (Subset=="P2unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Pig_feces_2")
} else if (Subset=="S1") {
  Metadata<-filter(Metadata, Experiment == "Sewage_1")
} else if (Subset=="S1spiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_1_spiked")
} else if (Subset=="S1unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_1")
} else if (Subset=="S2") {
  Metadata<-filter(Metadata, Experiment == "Sewage_2")
} else if (Subset=="S2spiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_2_spiked")
} else if (Subset=="S2unspiked") {
  Metadata<-filter(Metadata, Sample_type == "Sewage_2")
} else if (Subset=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubExp=="HX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata<-filter(Metadata, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata<-filter(Metadata, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata2<-filter(Metadata, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata2$Matching_samples)
  Metadata<-filter(Metadata, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata2)
} else if (SubExp=="HXFTX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata<-filter(Metadata, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubFre=="Frozen") {
  Metadata<-filter(Metadata, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata<-filter(Metadata, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

##Wanted to test if we for instance removed the 22C samples
#RemoveConditions<-c("64h22C", "16h22C")
#Metadata<-Metadata[!Metadata$Replicate_Boxplot %in% RemoveConditions, ]


#Applying subsetting to OTU tables 
Tax<-select(Tax, one_of(Metadata$Sample))
```

### 3 Long formatted Bray Curtis dissimilarities
```{r}
#Multi dimensional scaling 
distmatrix <- vegdist(decostand(t(Tax), method=Stand), method=Dist)

#Make distances into matrix
distmatrix<-as.matrix(distmatrix)

#Contains all values except the ones equal to 0
meltPwBC<-subset(melt(distmatrix), value!=0)
rm(distmatrix)

#Have comparisons of both 1vs2 and 2vs1 ..., but are removed during the following filtration
```

### 4 Boxplot of dissimilarities
```{r, echo=FALSE}
###############To make: 2 All between between sample groups#################
##Only contain samples with "P1" in column 1 named Var1
P1vsOther<-filter(meltPwBC, grepl('P1', Var1))
##From this exclude comparisons with samples containing "P1"
P1vsOther<-filter(P1vsOther, !grepl('P1', Var2))
##Histogram to look at distribution
hist(P1vsOther$value)
###Low values are the other fecal samples the high values are the sewage samples

##Only contain samples with "P2" in column 1 named Var1
P2vsS1and2<-filter(meltPwBC, grepl('P2', Var1))
##From this exclude comparisons with samples containing "P2" and "P1" because that comparison already is made
P2vsS1and2<-filter(P2vsS1and2, !grepl('P1|P2', Var2))

##Only contain samples with "S1" in column 1 named Var1
S1vsS2<-filter(meltPwBC, grepl('S1', Var1))
##and "S2" in column 2
S1vsS2<-filter(S1vsS2, grepl('S2', Var2))

#Combine all pairwise comparisons
Allbetweensamplegroups<-bind_rows(P1vsOther, P2vsS1and2, S1vsS2)

#Create boxplot 2
boxplot(Allbetweensamplegroups$value)

#Structure dataframe for full boxplot 
box2<-mutate(Allbetweensamplegroups, name="Allbetween")


###############To make: 3 Feces 1 vs. Feces 2##############
##Only contain samples with "P1" in column 1 named Var1
P1vsP2<-filter(meltPwBC, grepl('P1', Var1))
##and "P2" in column 2 
P1vsP2<-filter(P1vsP2, grepl('P2', Var2))
##Histogram to look at distribution
hist(P1vsP2$value)
###
##Create boxplot
boxplot(P1vsP2$value)

#Structure dataframe for full boxplot 
box3<-mutate(P1vsP2, name="P1vsP2")



###############To make: 4 Feces 1 vs. Sewage 1##############
##Only contain samples with "P1" in column 1 named Var1
P1vsS1<-filter(meltPwBC, grepl('P1', Var1))
##and "S1" in column 2 
P1vsS1<-filter(P1vsS1, grepl('S1', Var2))
##Histogram to look at distribution
hist(P1vsS1$value)
###
##Create boxplot
boxplot(P1vsS1$value)

#Structure dataframe for full boxplot 
box4<-mutate(P1vsS1, name="P1vsS1")

###############To make: 5 Feces 1 vs. Sewage 2##############
##Only contain samples with "P1" in column 1 named Var1
P1vsS2<-filter(meltPwBC, grepl('P1', Var1))
##and "S2" in column 2 
P1vsS2<-filter(P1vsS2, grepl('S2', Var2))
##Histogram to look at distribution
hist(P1vsS2$value)
###
##Create boxplot
boxplot(P1vsS2$value)

#Structure dataframe for full boxplot 
box5<-mutate(P1vsS2, name="P1vsS2")

###############To make: 6 Feces 2 vs. Sewage 1##############
##Only contain samples with "P2" in column 1 named Var1
P2vsS1<-filter(meltPwBC, grepl('P2', Var1))
##and "S1" in column 2 
P2vsS1<-filter(P2vsS1, grepl('S1', Var2))
##Histogram to look at distribution
hist(P2vsS1$value)
###
##Create boxplot
boxplot(P2vsS1$value)

#Structure dataframe for full boxplot 
box6<-mutate(P2vsS1, name="P2vsS1")

###############To make: 7 Feces 2 vs. Sewage 2##############
##Only contain samples with "P2" in column 1 named Var1
P2vsS2<-filter(meltPwBC, grepl('P2', Var1))
##and "S2" in column 2 
P2vsS2<-filter(P2vsS2, grepl('S2', Var2))
##Histogram to look at distribution
hist(P2vsS2$value)
###
##Create boxplot
boxplot(P2vsS2$value)

#Structure dataframe for full boxplot 
box7<-mutate(P2vsS2, name="P2vsS2")

###############To make: 8 Sewage 1 vs. Sewage 2##############
##Only contain samples with "S1" in column 1 named Var1
S1vsS2<-filter(meltPwBC, grepl('S1', Var1))
##and "S2" in column 2 
S1vsS2<-filter(S1vsS2, grepl('S2', Var2))
##Histogram to look at distribution
hist(S1vsS2$value)
###
##Create boxplot
boxplot(S1vsS2$value)

#Structure dataframe for full boxplot 
box8<-mutate(S1vsS2, name="S1vsS2")

###############To make: 1 All within sample groups (Also 9-12)##############
##Only contain samples with "P1" in column 1 named Var1
P1vsP1<-filter(meltPwBC, grepl('P1', Var1))
##and "P1" in column 2 
P1vsP1<-filter(P1vsP1, grepl('P1', Var2))
##Histogram to look at distribution
hist(P1vsP1$value)
###
##Create boxplot 9 P1 vs P1
boxplot(P1vsP1$value)

#Structure dataframe for full boxplot 
box9<-mutate(P1vsP1, name="P1vsP1")

##Only contain samples with "P2" in column 1 named Var1
P2vsP2<-filter(meltPwBC, grepl('P2', Var1))
##and "P2" in column 2 
P2vsP2<-filter(P2vsP2, grepl('P2', Var2))
##Histogram to look at distribution
hist(P2vsP2$value)
###
##Create boxplot 10 P2 vs P2
boxplot(P2vsP2$value)

#Structure dataframe for full boxplot 
box10<-mutate(P2vsP2, name="P2vsP2")

##Only contain samples with "S1" in column 1 named Var1
S1vsS1<-filter(meltPwBC, grepl('S1', Var1))
##and "S1" in column 2 
S1vsS1<-filter(S1vsS1, grepl('S1', Var2))
##Histogram to look at distribution
hist(S1vsS1$value)
###
##Create boxplot 11 S1 vs S1
boxplot(S1vsS1$value)

#Structure dataframe for full boxplot 
box11<-mutate(S1vsS1, name="S1vsS1")

##Only contain samples with "S2" in column 1 named Var1
S2vsS2<-filter(meltPwBC, grepl('S2', Var1))
##and "S2" in column 2 
S2vsS2<-filter(S2vsS2, grepl('S2', Var2))
##Histogram to look at distribution
hist(S2vsS2$value)
###
##Create boxplot 12 S2 vs S2
boxplot(S2vsS2$value)

#Structure dataframe for full boxplot 
box12<-mutate(S2vsS2, name="S2vsS2")


#Combine all pairwise comparisons of the sample samples
Allwithinsamplegroups<-bind_rows(P1vsP1, P2vsP2, S1vsS1, S2vsS2)

#Create boxplot 2
boxplot(Allwithinsamplegroups$value)

#Structure dataframe for full boxplot 
box1<-mutate(Allwithinsamplegroups, name="Allwithin")
```

### 5 Boxplot containing all
```{r}
boxplotting<-bind_rows(box1, box2, box3, box4, box5, box6, box7, box8, box9, box10, box11, box12)

#Defining the order, can change if needed
boxplotting$name2<-factor(boxplotting$name, levels= c("Allbetween", "Allwithin", "P1vsP1", "P2vsP2", "S1vsS1", "S2vsS2", "P1vsP2", "S1vsS2", "P1vsS1", "P1vsS2", "P2vsS1", "P2vsS2"))

#png(paste("BoxplotBCstat", "Genus", ".png", sep=""))
ggplot(boxplotting, aes(x=name2, y=value))+geom_boxplot()+ggtitle(paste("Boxplot", "Genus", sep=" ")) + labs(x="Compared groups", y="Bray-Curtis dissimilarities \n calculated on Hellinger transformed data")
#dev.off()
```

### 6 Adding Replicates to boxplot figure
```{r}
#Makes column based on Var1 that only contain middle part of sample name
boxplotting2 <- mutate(boxplotting, from = str_replace(Var1,"[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_","") %>% str_replace("_[a-z]$",""))
#Makes column based on Var2 that only contain middle part of sample name
boxplotting2 <- mutate(boxplotting2, to = str_replace(Var2,"[A-X0-9]*_[A-X0-9]*_[A-X0-9]*_","") %>% str_replace("_[a-z]$",""))
#Makes a column were TRUE means from is the same as to otherwise FALSE
boxplotting2 <- mutate(boxplotting2, rep = (from == to))

##Only contain samples that are replicates containing TRUE in rep
Replicates<-filter(boxplotting2, grepl('TRUE', rep))
#Now both represented in all within and P1vsP1...
Replicates<-filter(Replicates, grepl('Allwithin', name))
#Now only represented once
boxplot(Replicates$value)

#Boxplot with all
#Structure dataframe for full boxplot 
Replicates<-mutate(Replicates, name="Replicates")
Replicates<-mutate(Replicates, name2="Replicates")

boxplotting2<-bind_rows(Replicates, boxplotting2)

#Defining the order, can change if needed
boxplotting2$name2<-factor(boxplotting2$name, levels= c("Replicates", "Allbetween", "Allwithin", "P1vsP1", "P2vsP2", "P1vsP2", "S1vsS1", "S2vsS2", "S1vsS2", "P1vsS1", "P1vsS2", "P2vsS1", "P2vsS2"))

pdf(paste("BoxplotBCstat", "Genus", SubExp, Subset, "2.pdf", sep=""), height=6, width=12)
ggplot(boxplotting2, aes(x=name2, y=value)) + geom_boxplot() + labs(x="Compared sample groups", y="Bray-Curtis dissimilarities \n calculated on Hellinger transformed data") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size=12, angle = 45, hjust = 1), axis.title=element_text(size=14)) 
dev.off()

#Trying to make fancy boxplots + geom_jitter(), might want to  
#pdf(paste("BoxplotBCstatFancy", "Genus", "2.pdf", sep=""), height=6, width=12)
ggplot(boxplotting2, aes(x=name2, y=value)) + geom_boxplot() + labs(x="Compared sample groups", y="Bray-Curtis dissimilarities \n calculated on Hellinger transformed data") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), axis.text.x = element_text(size=12, angle = 45, hjust = 1), axis.title=element_text(size=14)) + geom_jitter() 
#,dev.off()
```


### 7 Statistical testing of differences between groups (parametric)
Make statistical test ANOVA and follow-up tests (t-tests)   
See below for non-parametric Kruskal-wallis and Mann-Whitney.
```{r}
#Model fitting
distmodel1<-lm(value~name, data=boxplotting)
summary(distmodel1)

distmodel1object<-anova(distmodel1)

if (distmodel1object[1,5]>0.05) {
  print("Anova not significant")
} else if (PostTestParametric=="ttest") {
  pairwise.t.test(boxplotting$value, boxplotting$name, p.adj=padj)
} else if (PostTestParametric=="TukeyHSD") {
  TukeyHSD(aov(value~name, data=boxplotting), "name")
} else {
  print("No valid post test specified")
}

#Using the TukeyHSD also incorporates correction
```

### 8 Levene's test
Levene's test, assessing if an ANOVA is appropriate. Can log transform data otherwise a non-parametric test such as Kruskal-Wallis followed by a Mann-Whitney test.    
Levene's test explained: If p-value is less than 0.05 the obtained differences in sample variances are unlikely to have occurred based on random sampling from a population with equal variances
```{r}
leveneTest(value~name, data=boxplotting)
leveneTest(value~name, data=boxplotting, center=mean)

#Removing Allbetween and Allwithin
Levene1<-filter(boxplotting, !grepl('Allbetween|Allwithin', name))
leveneTest(value~name, data=Levene1)
leveneTest(value~name, data=Levene1, center=mean)

#Removing all within comparisons
Levene2<-filter(boxplotting, grepl('P1vsP2|S1vsS2|P1vsS1|P1vsS2|P2vsS1|P2vsS2', name))
leveneTest(value~name, data=Levene2)
leveneTest(value~name, data=Levene2, center=mean)


#Levenes test on log transformed values
boxplotting$logvalue<-log(boxplotting$value)

leveneTest(logvalue~name, data=boxplotting)
leveneTest(logvalue~name, data=boxplotting, center=mean)

#Removing Allbetween and Allwithin
Levene1<-filter(boxplotting, !grepl('Allbetween|Allwithin', name))
leveneTest(logvalue~name, data=Levene1)
leveneTest(logvalue~name, data=Levene1, center=mean)

#Removing all within comparisons
Levene2<-filter(boxplotting, grepl('P1vsP2|S1vsS2|P1vsS1|P1vsS2|P2vsS1|P2vsS2', name))
leveneTest(logvalue~name, data=Levene2)
leveneTest(logvalue~name, data=Levene2, center=mean)
```

The Levene's test is significant it seem appropiate to make a non-parametric test to test if there is a difference between the groups.

### 9 Statistical testing of differences between groups (non-parametric)
Make statistical test Kruskal-wallis and follow-up tests (Mann-Whitney).
```{r}
kruskal<-kruskal.test(boxplotting$value~as.factor(boxplotting$name))
kruskal #Just copy output
#write.table(kruskal, file=paste("Kruskal", Subset, ".txt", sep=""), sep="\t", dec=".", row.names = T, quote = F)

if (kruskal$p.value>0.05) {
  print("Kruskal test not significant")
} else if (PostTestNonParametric=="Dunne") {
  dunnTest(boxplotting$value~as.factor(boxplotting$name), method=Dunnepadj)
} else if (PostTestNonParametric=="MannWhitney") {
  pairwise.wilcox.test(boxplotting$value, as.factor(boxplotting$name), p.adjust.method=MWpadj)
} else {
  print("No valid post test specified")
}

dunn<-dunnTest(boxplotting$value~as.factor(boxplotting$name), method=Dunnepadj)
dunn$res
write.table(dunn$res, file=paste("Pvalues", Subset, PostTestNonParametric, Dunnepadj, ".txt", sep=""), sep="\t", dec=".", row.names = T, quote = F)
```

## Comments
Follow-up test pairwise Mann-Whitney, Dunne .... From rcompanion.org/rcompanion/d_06.html "If the Kruskal-Wallis test is significant, a post-hoc analysis can be performed to determine which levels of the independent variable differ from each other level.  Probably the most popular test for this is the Dunn test, which is performed with the dunnTest function in the FSA package.  Adjustments to the p-values could be made using the method option to control the familywise error rate or to control the false discovery rate.  See ?p.adjust for details.  
Zar (2010) states that the Dunn test is appropriate for groups with unequal numbers of observations.  
Dunnne seems like an appropriate follow-up test since there are unequal numbers of observations also make Mann-Whitney as option


## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```

