---
title: "CCA"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
Creating CCA plots with subsetting to sample level and if samples were spiked or unspiked.    
Validation plots in the form of stressplots and screeplots are included.  
Colouring of genera were performed according to the domains of life and within bacteria the phyla containing most genera, these were also the most abundant phyla, finally the class Negativicutes within Firmicutes were categorized by themselves.

### Packages
```{r}
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2") 
library(ggthemes) #This package contains extra themes, scales, and geoms, and functions for and related to ggplot2 eg. scale_colour_gdocs() help(package="ggthemes")
library(plotly) #For making interactive web graphics. help(package="plotly") 
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr")
library(gridExtra)
library(reshape2)
library(tidyr)
library(cowplot)
```

## Define variables
Subset is a vector that is used to loop across the defined subsetting
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<- c("P1spiked", "P1unspiked", "P2spiked", "P2unspiked", "S1spiked", "S1unspiked", "S2spiked", "S2unspiked") 
#c("P1spiked", "P1unspiked", "P2spiked", "P2unspiked", "S1spiked", "S1unspiked", "S2spiked", "S2unspiked") #All, Allspiked, Allunspiked, PF, SW, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Filtering which organisms to include
OrgFlt<-"All" #All, Bac, BacSacCry, BacArcVirSacCry 

#Filtering of the count table. This is especially important to remove organisms after subsetting that are for instance not represented in pig feces 1 spiked ..., In the MAplot you can also see that the removed organisms are the low abundant ones that are already shrunk
Filter_Rows<-1 #Is based on row sums so can take integer values 
```

## Analysis
### 1 Read in data
```{r}
Tax <- read.delim(file="TaxonomyRaw20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Feature <- read.delim(file="FeatureShort20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```

### 2 Subsetting and creating CCAs stored in lists
```{r}
# Create a list to hold the plot objects.
ScreeList <- list()
StressList <- list()
CCAList <- list()

for (i in Subset) {
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting Metadata2
if (i=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (i=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (i=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (i=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (i=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (i=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (i=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (i=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (i=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (i=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (i=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (i=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (i=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (i=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (i=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (i=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (i=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
Tax2<-select(Tax, one_of(Metadata2$Sample))


#Filtering of the Counttable depending on rowSums. 
Tax2 <- Tax2[rowSums(Tax2)>Filter_Rows*length(Tax2),]

#Add title to plots
if (i=="P1spiked") {
  P<-"CCA pig feces 1 spiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="P1unspiked") {
  P<-"CCA pig feces 1 unspiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="P2spiked") {
  P<-"CCA pig feces 2 spiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="P2unspiked") {
  P<-"CCA pig feces 2 unspiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="S1spiked") {
  P<-"CCA sewage 1 spiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="S1unspiked") {
  P<-"CCA sewage 1 unspiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="S2spiked") {
  P<-"CCA sewage 2 spiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else if (i=="S2unspiked") {
  P<-"CCA sewage 2 unspiked"
  C<-"Screeplot"
  S<-"Stressplot"
} else {
  print("Subset defined not valid")
}


#Feature is merged there the spiked in orgs will be removed as well

#Norm of Tax. Should be taken into account. The chi square distance that is the basis of cca makes raw counts appropriate. See https://sites.google.com/site/mb3gustame/reference/dissimilarity 

#cca
cca <- cca(t(Tax2)~Replicate_Boxplot, Metadata2)

#Screeplot 
screeplot<-data.frame(cca$CCA$eig)
colnames(screeplot)<-c("eig")
screeplot$eig <- screeplot$eig[1:length(screeplot$eig)] / sum(screeplot$eig) * 100
screeplot<-add_rownames(screeplot, "MDS")
#screeplot$MDS <- factor(screeplot$MDS, levels=c(sprintf("MDS%d", 1:length(screeplot$eig))))

#Create plot name
pltName <- paste( 'scree', i, sep = '' )
#create screeplot
ScreeList[[ pltName ]]<-ggplot(screeplot, aes(x=MDS, y=eig)) + 
  geom_bar(stat="identity") + 
  labs(x ="MDS", y ="eig (%)") + 
  ggtitle(C) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12), axis.text.x=element_blank(), axis.ticks.x=element_blank()) 
#ggsave(filename=paste("ScreeplotCCA", "Genus", i, OrgFlt, ".pdf", sep=""), height=6, width=12)


#Create stressplot in ggplot source code: https://github.com/vegandevs/vegan/blob/master/R/stressplot.wcmdscale.R#L56, go to 'stressplot.cca'
k<-2
u <- cbind(cca$CCA$u, cca$CA$u)
sev <- sqrt(c(cca$CCA$eig, cca$CA$eig))
w <- sqrt(cca$rowsum)
u <- diag(w) %*% u %*% diag(sev, length(sev))
v <- cbind(cca$CCA$v, cca$CA$v)
v <- diag(sqrt(cca$colsum)) %*% v
## check that k <= rank
if (k > length(sev)) {
     warning(gettextf("max allowed rank is k = %d", length(sev)))
     k <- min(k, length(sev))
    }   
## Distances
Xbar <- u %*% t(v)
Xbark <- u[,seq_len(k), drop = FALSE] %*% t(v[,seq_len(k), drop = FALSE])
if (!is.null(cca$pCCA)) {
    pFit <- ordiYbar(cca, "pCCA")
    Xbar <- Xbar + pFit
    Xbark <- Xbark + pFit
}
dis <- dist(Xbar)
df <- melt(as.matrix(dis))
odis <- dist(Xbark)
df2 <- melt(as.matrix(odis))

#Create plot name
pltName <- paste( 'stress', i, sep = '' )
#create stressplot
StressList[[ pltName ]]<-ggplot() + 
  geom_point(aes(df$value, df2$value)) +
  ggtitle(S) + 
  labs(x = "Observed dissimilarity", y = "Ordination distance") + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12))



#Extract scores from cca object
scores <- vegan::scores(cca, display=c("sp", "wa", "cn", "bp"), choices=c(1,2,3))

#Extract site information and add to metadata
sites <- data.frame(scores$sites)
sites$Sample <- rownames(sites)
Metadata2 <- merge(Metadata2, sites, by="Sample")
#Creating column in metadata for plotting with lines between replicates
Metadata2$Sample_name_norep<-gsub("*_a|*_b|*_c", "", Metadata2$Sample_name)

#Extract org information and add to Feature data
species <- data.frame(scores$species)
species$Genus <- rownames(species)
Feature2 <- merge(Feature, species, by="Genus")
#Create column org grouping for colouring 
Feature2$col <- ifelse(Feature2$Phylum=="Firmicutes", "Firmicutes",
          ifelse(Feature2$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(Feature2$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(Feature2$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(Feature2$Domain=="Eukaryota", "Eukaryota",
          ifelse(Feature2$Domain=="Viruses", "Viruses",
          ifelse(Feature2$Domain=="Archaea", "Archaea",
                 "Other")))))))
Feature2$col[Feature2$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
Feature2$col <- factor(Feature2$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))

#Extract information on inertia 
inertia <- cca$CCA$tot.chi/cca$tot.chi
subheader <- paste("Inertia constrained by the explanatory variables", round(inertia, digits=2))
#Extract information on eig
eig1 <- (cca$CCA$eig[1]/(sum(cca$CCA$eig)+sum(cca$CA$eig)))*100
eig_1 <- paste("CCA1", round(eig1, digits=2), "% of total variance")
eig2 <- (cca$CCA$eig[2]/(sum(cca$CCA$eig)+sum(cca$CA$eig)))*100
eig_2 <- paste("CCA2", round(eig2, digits=2), "% of total variance")

#Change order in Replicate_Boxplot
#Metadata2$Replicate_Boxplot<-ordered(Metadata2$Replicate_Boxplot, levels=c("0h", "16h-80C", "64h-80C", "16h-20C", "64h-20C", "16h5C", "64h5C", "16h22C", "64h22C"))
#Change temperature and time to characters for plotting
Metadata2$Time<-as.character(Metadata2$Time)


#Example of how to make custom colors for the sample points changing both color and fill to represent the color schemes defined previously. 
#This is for P1, there are small differences in the number of samples in each group you need to be aware of
if (i=="P1spiked" || i=="P2spiked" || i=="P2unspiked" || i=="S1unspiked" || i=="S2spiked") {
#Create plot name
pltName <- paste('CCA', i, sep = '' )
CCAList[[ pltName ]] <- ggplot() + 
  geom_point(data=Metadata2[1:3,], aes(x=CCA1, y=CCA2, shape=Time), color="#999999", size=1.5, fill="#999999", stroke=1.2) + 
  geom_point(data=Metadata2[4:6,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[7:9,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[10:12,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[13:15,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_point(data=Metadata2[16:18,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[19:21,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[22:24,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[25:27,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_line(data=Metadata2, aes(x=CCA1, y=CCA2, group=Replicate_Boxplot)) + 
  geom_point(data=Feature2, aes(x=CCA1, y=CCA2, color=col, group=Genus), size=1) + 
  scale_colour_gdocs() + 
  labs(colour="Organisms", shape="Storage conditions", title=P, subtitle=subheader, x=eig_1, y=eig_2) + 
  scale_shape_manual(values=c(21, 22, 23)) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12), legend.position="none")
} else if (i=="P1unspiked") {
  #Create plot name
pltName <- paste('CCA', i, sep = '' )
CCAList[[ pltName ]] <- ggplot() + 
  geom_point(data=Metadata2[1:3,], aes(x=CCA1, y=CCA2, shape=Time), color="#999999", size=1.5, fill="#999999", stroke=1.2) + 
  geom_point(data=Metadata2[4:6,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[7:9,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[10:12,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[13:14,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_point(data=Metadata2[15:17,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[18:20,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[21:23,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[24:26,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_line(data=Metadata2, aes(x=CCA1, y=CCA2, group=Replicate_Boxplot)) + 
  geom_point(data=Feature2, aes(x=CCA1, y=CCA2, color=col, group=Genus), size=1) + 
  scale_colour_gdocs() + 
  labs(colour="Organisms", shape="Storage conditions", title=P, subtitle=subheader, x=eig_1, y=eig_2) + 
  scale_shape_manual(values=c(21, 22, 23)) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12), legend.position="none")
} else if (i=="S1spiked") {
  #Create plot name
pltName <- paste('CCA', i, sep = '' )
CCAList[[ pltName ]] <- ggplot() + 
  geom_point(data=Metadata2[1:3,], aes(x=CCA1, y=CCA2, shape=Time), color="#999999", size=1.5, fill="#999999", stroke=1.2) + 
  geom_point(data=Metadata2[4:6,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[7:9,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[10:12,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[13:15,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_point(data=Metadata2[16:18,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[19:20,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[21:23,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[24:25,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_line(data=Metadata2, aes(x=CCA1, y=CCA2, group=Replicate_Boxplot)) + 
  geom_point(data=Feature2, aes(x=CCA1, y=CCA2, color=col, group=Genus), size=1) + 
  scale_colour_gdocs() + 
  labs(colour="Organisms", shape="Storage conditions", title=P, subtitle=subheader, x=eig_1, y=eig_2) + 
  scale_shape_manual(values=c(21, 22, 23)) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12), legend.position="none")
} else if (i=="S2unspiked") {
  #Create plot name
pltName <- paste('CCA', i, sep = '' )
CCAList[[ pltName ]] <- ggplot() + 
  geom_point(data=Metadata2[1:3,], aes(x=CCA1, y=CCA2, shape=Time), color="#999999", size=1.5, fill="#999999", stroke=1.2) + 
  geom_point(data=Metadata2[4:5,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[6:8,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[9:11,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[12:14,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_point(data=Metadata2[15:17,], aes(x=CCA1, y=CCA2, shape=Time), color="#0571b0", size=1.5, fill="#0571b0", stroke=1.2) + 
  geom_point(data=Metadata2[18:20,], aes(x=CCA1, y=CCA2, shape=Time), color="#92c5de", size=1.5, fill="#92c5de", stroke=1.2) + 
  geom_point(data=Metadata2[21:23,], aes(x=CCA1, y=CCA2, shape=Time), color="#f4a582", size=1.5, fill="#f4a582", stroke=1.2) + 
  geom_point(data=Metadata2[24:26,], aes(x=CCA1, y=CCA2, shape=Time), color="#ca0020", size=1.5, fill="#ca0020", stroke=1.2) + 
  geom_line(data=Metadata2, aes(x=CCA1, y=CCA2, group=Replicate_Boxplot)) + 
  geom_point(data=Feature2, aes(x=CCA1, y=CCA2, color=col, group=Genus), size=1) + 
  scale_colour_gdocs() + 
  labs(colour="Organisms", shape="Storage conditions", title=P, subtitle=subheader, x=eig_1, y=eig_2) + 
  scale_shape_manual(values=c(21, 22, 23)) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12), legend.position="none")
  } else {
  print("Error in i")
}

}
```

### 3 Create figures 
```{r}
#Have the plots stored in lists
lay <- rbind(c(1,1,4,4,7,7,10,10),
             c(1,1,4,4,7,7,10,10),
             c(2,3,5,6,8,9,11,12),
             c(13,13,16,16,19,19,22,22),
             c(13,13,16,16,19,19,22,22),
             c(14,15,17,18,20,21,23,24))
pdf(paste("CCA", "Genus", ".pdf", sep=""), width=24, height=12)
grid.arrange(CCAList$CCAP1unspiked, ScreeList$screeP1unspiked, StressList$stressP1unspiked, CCAList$CCAP2unspiked, ScreeList$screeP2unspiked, StressList$stressP2unspiked, CCAList$CCAS1unspiked, ScreeList$screeS1unspiked, StressList$stressS1unspiked, CCAList$CCAS2unspiked, ScreeList$screeS2unspiked, StressList$stressS2unspiked, CCAList$CCAP1spiked, ScreeList$screeP1spiked, StressList$stressP1spiked, CCAList$CCAP2spiked, ScreeList$screeP2spiked, StressList$stressP2spiked, CCAList$CCAS1spiked, ScreeList$screeS1spiked, StressList$stressS1spiked, CCAList$CCAS2spiked, ScreeList$screeS2spiked, StressList$stressS2spiked, layout_matrix = lay)
dev.off()

#Alternative lay
#Have the plots stored in lists
lay <- rbind(c(1,1,13,13),
             c(1,1,13,13),
             c(2,3,14,15),
             c(4,4,16,16),
             c(4,4,16,16),
             c(5,6,17,18),
             c(7,7,19,19),
             c(7,7,19,19),
             c(8,9,20,21),
             c(10,10,22,22),
             c(10,10,22,22),
             c(11,12,23,24))
pdf(paste("SF6CCA", "Tall", ".pdf", sep=""), width=12, height=20)
grid.arrange(CCAList$CCAP1unspiked, ScreeList$screeP1unspiked, StressList$stressP1unspiked, CCAList$CCAP2unspiked, ScreeList$screeP2unspiked, StressList$stressP2unspiked, CCAList$CCAS1unspiked, ScreeList$screeS1unspiked, StressList$stressS1unspiked, CCAList$CCAS2unspiked, ScreeList$screeS2unspiked, StressList$stressS2unspiked, CCAList$CCAP1spiked, ScreeList$screeP1spiked, StressList$stressP1spiked, CCAList$CCAP2spiked, ScreeList$screeP2spiked, StressList$stressP2spiked, CCAList$CCAS1spiked, ScreeList$screeS1spiked, StressList$stressS1spiked, CCAList$CCAS2spiked, ScreeList$screeS2spiked, StressList$stressS2spiked, layout_matrix = lay)
dev.off()

png(paste("SF6CCA", "Tall", ".png", sep=""), width=900, height=1500)
grid.arrange(CCAList$CCAP1unspiked, ScreeList$screeP1unspiked, StressList$stressP1unspiked, CCAList$CCAP2unspiked, ScreeList$screeP2unspiked, StressList$stressP2unspiked, CCAList$CCAS1unspiked, ScreeList$screeS1unspiked, StressList$stressS1unspiked, CCAList$CCAS2unspiked, ScreeList$screeS2unspiked, StressList$stressS2unspiked, CCAList$CCAP1spiked, ScreeList$screeP1spiked, StressList$stressP1spiked, CCAList$CCAP2spiked, ScreeList$screeP2spiked, StressList$stressP2spiked, CCAList$CCAS1spiked, ScreeList$screeS1spiked, StressList$stressS1spiked, CCAList$CCAS2spiked, ScreeList$screeS2spiked, StressList$stressS2spiked, layout_matrix = lay)
dev.off()


legend<-ggplot() + 
  geom_point(data=Metadata2[1:3,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[4:5,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[6:8,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[9:11,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[12:14,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[15:17,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[18:20,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[21:23,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_point(data=Metadata2[24:26,], aes(x=CCA1, y=CCA2, shape=Time), color="Black", size=1.5, fill="Black", stroke=1.2) + 
  geom_line(data=Metadata2, aes(x=CCA1, y=CCA2, group=Replicate_Boxplot)) + 
  geom_point(data=Feature2, aes(x=CCA1, y=CCA2, color=col, group=Genus), size=2) + 
  scale_colour_gdocs() + 
  labs(colour="Organisms", shape="Time", title=paste("CCA ", i), subtitle=subheader, x=eig_1, y=eig_2) + 
  scale_shape_manual(values=c(21, 22, 23)) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title=element_text(size=12))
legendplot<-get_legend(legend)
pdf(paste("LegendCCA", "Genus", ".pdf", sep=""), width=24, height=12)
grid.arrange(legendplot)
dev.off()
```


## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```