---
title: "DESeq results for creating MA plots and comparing across 1"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
The purpose of the following code is to create DESeq results that are compatible with "SF7to10DESeqResMAcompAcross2.Rmd".  
In the script MA plots are saved in a list and tables containing DESeq results.  
Have to change Submor for specific desired comparison and the Titleling 
Other variables that can be changed include Test_Factor and Filter_Rows

### Packages
```{r}
library(DESeq2) #Differential analysis of count data. help(package="DESeq2")
library(knitr) #A general-purpose package for dynamic report generation in R. help(package="knitr")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr")
library(pheatmap) # Implementation of heatmaps that offers more control over dimensions and appearance help(package="pheatmap")
library(ggplot2) #Data visualization, based on grammar of graphics. help(package="ggplot2")  
library(plotly) #For making interactive web graphics. help(package="plotly") 
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
library(ggthemes) #This package contains extra themes, scales, and geoms, and functions for and related to ggplot2 eg. scale_colour_gdocs() help(package="ggthemes")
```

## Define variables
Remember to change Submor according to desired analysis
```{r}
#Define normalization strategy 
##DESeq uses raw counts but I supply costum size factors

#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<-c("P1spiked", "P1unspiked", "P2spiked", "P2unspiked", "S1spiked", "S1unspiked", "S2spiked", "S2unspiked")
#c("P1spiked", "P1unspiked", "P2spiked", "P2unspiked", "S1spiked", "S1unspiked", "S2spiked", "S2unspiked")
#All, Allspiked, Allunspiked, PF, PFspiked, PFunspiked, SW, SWspiked, SWunspiked, P1, P1spiked, P1unspiked, P2, P2spiked, P2unspiked, S1, S1spiked, S1unspiked, S2, S2spiked, S2unspiked 

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Subset time temp dependant
SubMor<-"0hvs-80C64h" #0hvs22C64h, 0hvs22C16h, 0hvs5C64h, 0hvs5C16h, 0hvsfrozen, 0hvs-20C64h, 0hvs-80C64h, 0hvs-20C16h, 0hvs-80C16h, All

#Create title depending on Submor
if (SubMor=="0hvs-80C64h") {
  Titleling<-"-80°C for 64h"
} else if (SubMor=="0hvs22C64h") {
  Titleling<-"22°C for 64h"
} else if (SubMor=="0hvs22C16h") {
  Titleling<-"22°C for 16h"
} else if (SubMor=="0hvs5C64h") {
  Titleling<-"5°C for 64h"
} else if (SubMor=="0hvs5C16h") {
  Titleling<-"5°C for 16h"
} else if (SubMor=="0hvsfrozen") {
  Titleling<-"frozen"
} else if (SubMor=="0hvs-20C64h") {
  Titleling<-"-20°C for 64h"
} else if (SubMor=="0hvs-20C16h") {
  Titleling<-"-20°C for 16h"
} else if (SubMor=="0hvs-80C16h") {
  Titleling<-"-80°C for 16h"
} else if (SubMor=="All") {
  Titleling<-"All"
} else {
  print("SubMor defined not valid")
}

Titleling <- "-80°C for 64h"

#Which factor to test can pick metadata columns 
Test_Factor <- "Replicate_Boxplot" #FrozenUnfrozenSimple, Replicate_Groups, as.factor(Time)+StoragePlacementType, as.factor(Time)+StoragePlacementType+as.factor(Freeze_thaw_cycles), ...

#Filtering of the count table. This is especially important to remove organisms after subsetting that are for instance not represented in pig feces 1 spiked ..., In the MAplot you can also see that the removed organisms are the low abundant ones that are already shrunk
Filter_Rows<-10 #Is based on row sums so can take integer values 

#Vegdist dissimilarity method. For sample clustering
Dist<-"bray" #manhattan, euclidean, canberra, bray, kulczynski, jaccard, gower, altGower, morisita, horn, mountford, raup, binomial, chao, cao or mahalanobis

#Decostand standardization method. For sample clustering
Stand<-"hellinger" #total, max, freq, normalize, range, standardize, pa, chi.square, hellinger, log

#The organisms clustering
OrgCluster<-"correlation" #correlation, euclidean, maximum, manhattan, canberra, binary, minkowski 

#Threshold for FDR-adjusted p-values
padj_threshold <- 0.01  
```

### Coloring scheme
```{r}
#Define coloring schemes
#Storagecol<-c("Direct" = "#999999", "-80" = "#0571b0", "-20" = "#92c5de", "5" = "#f4a582", "22" = "#ca0020")
#Experimentcol<-c("Pig_feces_1" = "#30241E", "Pig_feces_2" = "#B59B80", "Sewage_1" = "#33a02c", "Sewage_2" = "#B2DF8A")
Orgcol<-c("Proteobacteria" = "#3366CC", "Bacteroidetes" = "#DC3912", "Actinobacteria" = "#FF9900", "Firmicutes" = "#109618", "Firmicutes(Negativicutes)" = "#990099", "Eukaryota" = "#0099C6", "Archaea" = "#DD4477", "Viruses" = "#66AA00", "Other" = "#B82E2E")

##How organisms were decided
#show_col(gdocs_pal()(9))
#gdocs_pal()(9) #to get the colors
#library(RColorBrewer)
#palette(brewer.pal(n=9, name="Set1")) 
#brewer.pal(n=9, name="Set1") #to get the colors overlaps with other specified colors

##How to specify one coloring scheme in ggplots using above
#scale_color_manual(values=Experimentcol) + 
##How to specify multiple colors in ggplots using above 
#scale_color_manual(values=c(Storagecol, Experimentcol)) + 
```

## Analysis
### 1 Read in data
```{r}
Tax <- read.delim(file="TaxonomyRaw20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Feature <- read.delim(file="FeatureShort20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```


### 2 Subsetting, DESeq and visualization 
Additional plots are created 
Using for loop to go through Subset vector
```{r}
MAList <- list()


for (i in Subset) {
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting Metadata2
if (i=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (i=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (i=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (i=="PFspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked" | Sample_type == "Pig_feces_2_spiked")
} else if (i=="PFunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1" | Sample_type == "Pig_feces_2")
} else if (i=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (i=="SWspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked" | Sample_type == "Sewage_2_spiked")
} else if (i=="SWunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1" | Sample_type == "Sewage_2")
} else if (i=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (i=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (i=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (i=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (i=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (i=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (i=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (i=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (i=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (i=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (i=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (i=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (i=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting metadata
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}


#Further subsetting metadata
if (SubMor=="0hvs22C64h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "64h22C")
} else if (SubMor=="0hvs22C16h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "16h22C")
} else if (SubMor=="0hvs5C64h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "64h5C")
} else if (SubMor=="0hvs5C16h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "16h5C")
} else if (SubMor=="0hvsfrozen") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "16h-20C"| Replicate_Boxplot == "64h-20C"| Replicate_Boxplot == "16h-80C"| Replicate_Boxplot == "64h-80C")
} else if (SubMor=="0hvs-20C64h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "64h-20C")
} else if (SubMor=="0hvs-80C64h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "64h-80C")
} else if (SubMor=="0hvs-20C16h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "16h-20C")
} else if (SubMor=="0hvs-80C16h") {
  Metadata2<-filter(Metadata2, Replicate_Boxplot == "0h" | Replicate_Boxplot == "16h-80C")
} else if (SubMor=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
Tax2<-select(Tax, one_of(Metadata2$Sample))


#Change Time to be a factor
#Metadata$Time<-as.factor(Metadata$Time)

###5.1 Setup DESeq2
######coldata is the metadata containing the columns batch and condition
######In design put the variable of interest at the end of the formula and make sure the control level is the first level

Metadata2$Replicate_Boxplot <- factor(Metadata2$Replicate_Boxplot, levels=c("0h", "16h22C", "64h22C", "16h5C", "64h5C", "16h-20C", "64h-20C", "16h-80C", "64h-80C"))

#Create design formula
design <- formula(paste("~ ", Test_Factor))

#Create DESeq2 object with matrix and metadata
dds <- DESeqDataSetFromMatrix(countData = Tax2,
                              colData = Metadata2,
                              design = design)

##See vignette for note on factor levels

#Filtering of the Counttable. Bare minimum is performed later. Also to reduce memory of the dds
dds <- dds[ rowSums(counts(dds))>Filter_Rows*length(Tax2),]

##Used when comparing the DESeq estimated size factors with Mapped parsed reads. 
#dds<-estimateSizeFactors(dds)
#DESeqSizeFactors<-sizeFactors(dds)
#SizeFactors<-as.data.frame(DESeqSizeFactors)

#Provide custom size factors
dds<-estimateSizeFactors(dds)
dds@colData@listData$sizeFactor
dds@colData@listData$sizeFactor <- colSums(Tax2)/mean(colSums(Tax2))
dds@colData@listData$sizeFactor

#Estimate dispersions and fit the GLM
dds <- DESeq(dds)






#https://bioconductor.org/packages/3.7/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#ma-plot

res<-results(dds)
summary(res)
resultsNames(dds)


#MA plots from inbuild DESeq function
plotMA(dds, alpha=0.1)
plotMA(res, ylim=c(-2,2))

#Data structuring
df <- as.data.frame(res)
df <- add_rownames(df, "Genus")
df <- merge(Feature, df, by="Genus") #Removes orgs that are not present in these samples
#Create column org grouping for colouring 
df$col <- ifelse(df$Phylum=="Firmicutes", "Firmicutes",
          ifelse(df$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(df$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(df$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(df$Domain=="Eukaryota", "Eukaryota",
          ifelse(df$Domain=="Viruses", "Viruses",
          ifelse(df$Domain=="Archaea", "Archaea",
                 "Other")))))))
df$col[df$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
df$col <- factor(df$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))
#Create column for sig level used for shape
df$shape <- ifelse(df$pvalue>0.1, "Not significant P>0.1",
            ifelse(df$pvalue<0.01, "Significant P<0.01",
                 "Borderline 0.1>P>0.01"))
df$shape <- factor(df$shape, levels=c("Not significant P>0.1", "Borderline 0.1>P>0.01", "Significant P<0.01"))

##Remove Cardiovirus in P2 0hvs5C64h & S1 0hvs22C64h
#df <- filter(df, Genus!="Cardiovirus")
##Remove Kurthia in P1 0hvs22C64h
#df <- filter(df, Genus!="Kurthia")

#Make MA plots ggplot
#Be aware that levels of variables change so what becomes positive/negative log2fold changes not anymore
pdf(paste("MAplot", "Genus", i, Test_Factor, Filter_Rows, ".pdf", sep=""), height=6, width=12)  
print(ggplot(df) + 
        geom_point(aes(x=baseMean, y=log2FoldChange, color=col, shape=shape, group=Genus), size=1) + 
        scale_x_log10() + 
        scale_color_manual(values=Orgcol) + 
        labs(title=paste("MA-plot", i, sep=" "), subtitle=paste("Positive log2 fold change indicate higher relative abundance in sample stored at", Titleling, "compared to immediate DNA-extraction")) + 
        theme_bw() + 
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #+ coord_cartesian(ylim=c(-5, 5)) #Can add to limit plotting boundaries + coord_cartesian(ylim=c(-5.5, 5)) #Remember levels of StoragePlacementType Nothing, Fridge, Table, Freezer, FreezerDeep. This is important for the specific comparison and what a positive log2foldchange represents. 
dev.off()


dfin <- filter(df, -2.5<log2FoldChange & log2FoldChange<2.5)
dfout <- filter(df, -2.5>log2FoldChange | 2.5<log2FoldChange)
dfout$log2FoldChange[dfout$log2FoldChange > 2.5] <- 2.5
dfout$log2FoldChange[dfout$log2FoldChange < (-2.5)] <- -2.5
pltName <- paste('MA', i, sep = '' )
MAList[[ pltName ]] <- ggplot() + 
  geom_point(data=dfin, aes(x=baseMean, y=log2FoldChange, color=col, group=Genus), size=1) + 
  geom_point(data=dfout, aes(x=baseMean, y=log2FoldChange, color=col, group=Genus), shape=17, size=1) + 
  scale_x_log10() + 
  scale_color_manual(values=Orgcol) + 
  labs(title=paste("MA-plot", i, sep=" ")) + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none") +
  coord_cartesian(ylim=c(-2.5, 2.5)) +
  geom_hline(yintercept = 0)
#+ coord_cartesian(ylim=c(-5, 5)) #Can add to limit plotting boundaries + coord_cartesian(ylim=c(-5.5, 5)) #Remember levels of StoragePlacementType Nothing, Fridge, Table, Freezer, FreezerDeep. This is important for the specific comparison and what a positive log2foldchange represents.

ggplotly(ggplot(df) + geom_point(aes(x=baseMean, y=log2FoldChange, color=col, group=Genus)) + scale_x_log10() + scale_colour_gdocs() + labs(title="MA-plot", subtitle=paste("Positive log2 fold change indicate higher relative abundance in sample stored at", Titleling, "compared to immediate DNA-extraction")) + geom_hline(yintercept = 0))

#Unshrunken/shrunken 
#res<-results(dds, addMLE=TRUE)
#resLFC <- DESeq2::lfcShrink(dds, coef=2)
#resLFC
#DESeq2::plotMA(resLFC, ylim=c(-2,2))

#Make histogram
pdf(paste("Hist", "Genus", i, Test_Factor, Filter_Rows, ".pdf", sep=""), height=6, width=12) 
hist(res$pvalue[res$baseMean > 1], breaks = 0:20/20,
     col = "grey50", border = "white")
dev.off()

#Show dispersion plot with shrinkage
pdf(paste("Dispersion_plot", Filter_Rows, "Genus", i, ".pdf", sep=""), height=9, width=12)
plotDispEsts(dds)
dev.off()

#Plot PCA
dds_rlog<-rlogTransformation(dds)
pdf(paste("PCA_Filter", Filter_Rows, "Genus", i, ".pdf", sep=""), height=9, width=9)
print(plotPCA(dds_rlog, intgroup=c("Replicate_Boxplot", "SpikedUnspiked"))+coord_fixed())
dev.off()

#Plot heatmap in the order of most variable organisms (n=30)
select=order(rowMeans(assay(dds_rlog)), decreasing=TRUE)[1:30]
#Plot heatmap with new clustering of Samples
distmatrix_Species <- vegdist(decostand(t(Tax2), method=Stand), method=Dist)
pheatmap( assay(dds_rlog)[select, ], 
          scale="row",
          annotation_col=as.data.frame(
            colData(dds_rlog)[, c("Replicate_Boxplot", "FrozenUnfrozenSimple")]), show_colnames = FALSE,
          clustering_distance_cols = distmatrix_Species,
          clustering_distance_rows = OrgCluster,
          filename=paste("Heatmap_Filter", Filter_Rows, "Genus", i, ".pdf", sep=""))

#Save table. 
write.table(df, file=paste("ResultsTable", "Genus", i, Test_Factor, Filter_Rows, ".txt", sep=""), sep="\t", dec=",", row.names = F, quote = F)
}
```

## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```
