---
title: "DESeq results for creating MA plots and comparing across 2"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
To run this you need to run "SF7to10DESeqResMAcompAcross1.Rmd" (don't clear environment before running). Transfer Resultstables generated in "SF7to10DESeqResMAcompAcross1.Rmd". Also need Featureshort in the same directory.  
This is not sensitive to the variable SubMor and no variables needs to be defined.

### Clear environment except MAList
```{r}
rm(list=setdiff(ls(), c("MAList", "Feature")))
```

### Packages
```{r}
library(reshape2) # Flexibly restructure and aggregate data using just two functions: melt and 'dcast' (or 'acast'). help(package="reshape2")
library(gridExtra)
library(grid)
library(lattice)
library(cowplot)
```

### Coloring scheme
```{r}
#Define coloring schemes
#Storagecol<-c("Direct" = "#999999", "-80" = "#0571b0", "-20" = "#92c5de", "5" = "#f4a582", "22" = "#ca0020")
#Experimentcol<-c("Pig_feces_1" = "#30241E", "Pig_feces_2" = "#B59B80", "Sewage_1" = "#33a02c", "Sewage_2" = "#B2DF8A")
Orgcol<-c("Proteobacteria" = "#3366CC", "Bacteroidetes" = "#DC3912", "Actinobacteria" = "#FF9900", "Firmicutes" = "#109618", "Firmicutes(Negativicutes)" = "#990099", "Eukaryota" = "#0099C6", "Archaea" = "#DD4477", "Viruses" = "#66AA00", "Other" = "#B82E2E")

##How organisms were decided
#show_col(gdocs_pal()(9))
#gdocs_pal()(9) #to get the colors
#library(RColorBrewer)
#palette(brewer.pal(n=9, name="Set1")) 
#brewer.pal(n=9, name="Set1") #to get the colors overlaps with other specified colors

##How to specify one coloring scheme in ggplots using above
#scale_color_manual(values=Experimentcol) + 
##How to specify multiple colors in ggplots using above 
#scale_color_manual(values=c(Storagecol, Experimentcol)) + 
```

## Analysis
### 1 Read in data and create dataframes unspiked
```{r}
#Read in data
P1 <- read.delim(file="ResultsTableGenusP1unspikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
#Add prefix to column names 
colnames(P1)[7:14]<-paste("P1", colnames(P1)[7:14], sep="_")
#Remove irrelevant taxonomy info
P1 <- P1[,c(1, 7:14)]

P2 <- read.delim(file="ResultsTableGenusP2unspikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
colnames(P2)[7:14]<-paste("P2", colnames(P2)[7:14], sep="_")
P2 <- P2[,c(1, 7:14)]

S1 <- read.delim(file="ResultsTableGenusS1unspikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
colnames(S1)[7:14]<-paste("S1", colnames(S1)[7:14], sep="_")
S1 <- S1[,c(1, 7:14)]

S2 <- read.delim(file="ResultsTableGenusS2unspikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
colnames(S2)[7:14]<-paste("S2", colnames(S2)[7:14], sep="_")
S2 <- S2[,c(1, 7:14)]
```

### 2 Create Boxplots unspiked 
Boxplots are comparing shared organisms in all unspiked samples 
```{r}
#Create df from orgs shared across all samples
SharedOrgs<-merge(P1, merge(P2, merge(S1, S2, by="Genus"), by="Genus"), by="Genus")
#Extract log2foldchanges
dfall <- data.frame(select(SharedOrgs, matches("Genus|log2FoldChange")))
#First column genus into rownames
dfall <- data.frame(dfall[,-1], row.names=dfall[,1])
#Transpose
dfall <- data.frame(t(dfall))
#Mean of sd 
meanOrgs<-round(mean(apply(dfall, 2, sd)), digits=3)
#Melt
newdfall<-melt(dfall)
#______
#Change colnames
colnames(newdfall) <- c("Genus", "value")
#Add feature info to merge with newdfall
#Feature <- read.delim(file="FeatureShort20180925.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
#Merge for plotting
dfplotting <- merge(newdfall, Feature, by="Genus")
#Create column org grouping for colouring 
dfplotting$col <- ifelse(dfplotting$Phylum=="Firmicutes", "Firmicutes",
          ifelse(dfplotting$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(dfplotting$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(dfplotting$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(dfplotting$Domain=="Eukaryota", "Eukaryota",
          ifelse(dfplotting$Domain=="Viruses", "Viruses",
          ifelse(dfplotting$Domain=="Archaea", "Archaea",
                 "Other")))))))
dfplotting$col[dfplotting$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
dfplotting$col <- factor(dfplotting$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))
#_______
#Make boxplots change order boxplots according to mean
pdf("UnspikedAllreorder.pdf", height=6, width=12) 
ggplot(dfplotting) + 
  geom_boxplot(aes(x=reorder(Genus, value, mean), y=value, color=col)) + 
  theme_bw() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  geom_hline(yintercept=0) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Boxplots log2Foldchanges", subtitle=paste("P1unspiked, P2unspiked, S1unspiked, S2unspiked, mean sd ", meanOrgs, sep=""), x="Genera", y="log2Foldchange") 
dev.off()
#Add legend.position="none" to theme when have solution regarding 

AllList <- list()
AllList[[ "UnspikedComp0hwX" ]]<-ggplot(dfplotting) + 
  geom_boxplot(aes(x=reorder(Genus, value, mean), y=value, color=col)) + 
  theme_bw() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none") + 
  geom_hline(yintercept=0) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Boxplots log2Foldchanges", subtitle=paste("P1unspiked, P2unspiked, S1unspiked, S2unspiked, mean sd ", meanOrgs, sep=""), x="Genera", y="log2Foldchange") 
```


### 3 Create scatterplots pig feces unspiked
Does not seem informative to make boxplots when there is only two measures instead made scatterplot of the two pig feces samples and two sewage samples seperately
```{r}
#Create df from orgs shared across all Pig fecal samples or Sewage
SharedOrgs<-merge(P1, P2, by="Genus")
#Extract log2foldchanges
dfall <- data.frame(select(SharedOrgs, matches("Genus|log2FoldChange")))
#______
#Merge for plotting
dfplotting <- merge(dfall, Feature, by="Genus")
#Create column org grouping for colouring 
dfplotting$col <- ifelse(dfplotting$Phylum=="Firmicutes", "Firmicutes",
          ifelse(dfplotting$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(dfplotting$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(dfplotting$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(dfplotting$Domain=="Eukaryota", "Eukaryota",
          ifelse(dfplotting$Domain=="Viruses", "Viruses",
          ifelse(dfplotting$Domain=="Archaea", "Archaea",
                 "Other")))))))
dfplotting$col[dfplotting$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
dfplotting$col <- factor(dfplotting$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))

#Linear model info
modelP<-lm(P2_log2FoldChange ~ P1_log2FoldChange, dfplotting)
#Model validation
plot(modelP)
#summary(modelP)
intercept = format(coef(modelP)[1], digits = 2)
slope = format(coef(modelP)[2], digits = 2)
rsq<-format(summary(modelP)$r.squared, digits=2)

#_______
#Plot
pdf("UnspikedP1P2.pdf", height=6, width=6) 
ggplot(dfplotting) + 
  geom_point(aes(x=P1_log2FoldChange, y=P2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Pig feces", subtitle=paste("P2 = ", slope, "*P1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange P1", y="log2FoldChange P2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=P1_log2FoldChange, y=P2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")
dev.off()

P1P2List <- list()
P1P2List[[ "UnspikedComp0hwX" ]]<-ggplot(dfplotting) + 
  geom_point(aes(x=P1_log2FoldChange, y=P2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Pig feces", x="log2FoldChange P1", y="log2FoldChange P2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=P1_log2FoldChange, y=P2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")

ggplotly(ggplot(dfplotting) + 
  geom_point(aes(x=P1_log2FoldChange, y=P2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Pig feces", subtitle=paste("P2 = ", slope, "*P1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange P1", y="log2FoldChange P2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=P1_log2FoldChange, y=P2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none"))
```

### 4 Create scatterplots sewage unspiked 
```{r}
#Create df from orgs shared across all Pig fecal samples or Sewage
SharedOrgs<-merge(S1, S2, by="Genus")
#Extract log2foldchanges
dfall <- data.frame(select(SharedOrgs, matches("Genus|log2FoldChange")))
#______
#Merge for plotting
dfplotting <- merge(dfall, Feature, by="Genus")
#Create column org grouping for colouring 
dfplotting$col <- ifelse(dfplotting$Phylum=="Firmicutes", "Firmicutes",
          ifelse(dfplotting$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(dfplotting$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(dfplotting$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(dfplotting$Domain=="Eukaryota", "Eukaryota",
          ifelse(dfplotting$Domain=="Viruses", "Viruses",
          ifelse(dfplotting$Domain=="Archaea", "Archaea",
                 "Other")))))))
dfplotting$col[dfplotting$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
dfplotting$col <- factor(dfplotting$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))

#Linear model info
modelS<-lm(S2_log2FoldChange ~ S1_log2FoldChange, dfplotting)
#Model validation
plot(modelS)
#summary(modelS)
intercept = format(coef(modelS)[1], digits = 2)
slope = format(coef(modelS)[2], digits = 2)
rsq<-format(summary(modelS)$r.squared, digits=2)

#_______
#Plot
pdf("UnspikedS1S2.pdf", height=6, width=6) 
ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", subtitle=paste("S2 = ", slope, "*S1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")
dev.off()

S1S2List <- list()
S1S2List[[ "UnspikedComp0hwX" ]]<-ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")


ggplotly(ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", subtitle=paste("S2 = ", slope, "*S1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none"))
```

### 5 Read in data and create dataframes spiked
The same analysis spiked
```{r}
#Read in data
P1 <- read.delim(file="ResultsTableGenusP1spikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
#Add prefix to column names 
colnames(P1)[7:14]<-paste("P1", colnames(P1)[7:14], sep="_")
#Remove irrelevant taxonomy info
P1 <- P1[,c(1, 7:14)]

P2 <- read.delim(file="ResultsTableGenusP2spikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
colnames(P2)[7:14]<-paste("P2", colnames(P2)[7:14], sep="_")
P2 <- P2[,c(1, 7:14)]

S1 <- read.delim(file="ResultsTableGenusS1spikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
colnames(S1)[7:14]<-paste("S1", colnames(S1)[7:14], sep="_")
S1 <- S1[,c(1, 7:14)]

S2 <- read.delim(file="ResultsTableGenusS2spikedReplicate_Boxplot10.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
colnames(S2)[7:14]<-paste("S2", colnames(S2)[7:14], sep="_")
S2 <- S2[,c(1, 7:14)]
```


### 6 Create Boxplots spiked 
Boxplots are comparing shared organisms in all spiked samples 
```{r}
#Create df from orgs shared across all samples
SharedOrgs<-merge(P1, merge(P2, merge(S1, S2, by="Genus"), by="Genus"), by="Genus")
#Extract log2foldchanges
dfall <- data.frame(select(SharedOrgs, matches("Genus|log2FoldChange")))
#First column genus into rownames
dfall <- data.frame(dfall[,-1], row.names=dfall[,1])
#Transpose
dfall <- data.frame(t(dfall))
#Mean of sd 
meanOrgs<-round(mean(apply(dfall, 2, sd)), digits=3)
#Melt
newdfall<-melt(dfall)
#______
#Change colnames
colnames(newdfall) <- c("Genus", "value")
#Add feature info to merge with newdfall
#Feature <- read.delim(file="FeatureShort20180925.txt", check.names=FALSE, stringsAsFactors = FALSE, strip.white=TRUE, dec=",")
#Merge for plotting
dfplotting <- merge(newdfall, Feature, by="Genus")
#Create column org grouping for colouring 
dfplotting$col <- ifelse(dfplotting$Phylum=="Firmicutes", "Firmicutes",
          ifelse(dfplotting$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(dfplotting$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(dfplotting$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(dfplotting$Domain=="Eukaryota", "Eukaryota",
          ifelse(dfplotting$Domain=="Viruses", "Viruses",
          ifelse(dfplotting$Domain=="Archaea", "Archaea",
                 "Other")))))))
dfplotting$col[dfplotting$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
dfplotting$col <- factor(dfplotting$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))
#_______
#Make boxplots change order boxplots according to mean
pdf("SpikedAllreorder.pdf", height=6, width=12) 
ggplot(dfplotting) + 
  geom_boxplot(aes(x=reorder(Genus, value, mean), y=value, color=col)) + 
  theme_bw() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  geom_hline(yintercept=0) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Boxplots log2Foldchanges", subtitle=paste("P1unspiked, P2unspiked, S1unspiked, S2unspiked, mean sd ", meanOrgs, sep=""), x="Genera", y="log2Foldchange") 
dev.off()
#Add legend.position="none" to theme when have solution regarding 

AllList[[ "SpikedP0hvs5C64h" ]]<-ggplot(dfplotting) + 
  geom_boxplot(aes(x=reorder(Genus, value, mean), y=value, color=col)) + 
  theme_bw() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none") + 
  geom_hline(yintercept=0) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Boxplots log2Foldchanges", subtitle=paste("P1unspiked, P2unspiked, S1unspiked, S2unspiked, mean sd ", meanOrgs, sep=""), x="Genera", y="log2Foldchange") 
```


### 7 Create scatterplots pig feces spiked
Does not seem informative to make boxplots when there is only two measures instead made scatterplot of the two pig feces samples and two sewage samples seperately
```{r}
#Create df from orgs shared across all Pig fecal samples or Sewage
SharedOrgs<-merge(P1, P2, by="Genus")
#Extract log2foldchanges
dfall <- data.frame(select(SharedOrgs, matches("Genus|log2FoldChange")))
#______
#Merge for plotting
dfplotting <- merge(dfall, Feature, by="Genus")
#Create column org grouping for colouring 
dfplotting$col <- ifelse(dfplotting$Phylum=="Firmicutes", "Firmicutes",
          ifelse(dfplotting$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(dfplotting$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(dfplotting$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(dfplotting$Domain=="Eukaryota", "Eukaryota",
          ifelse(dfplotting$Domain=="Viruses", "Viruses",
          ifelse(dfplotting$Domain=="Archaea", "Archaea",
                 "Other")))))))
dfplotting$col[dfplotting$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
dfplotting$col <- factor(dfplotting$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))

#Linear model info
modelP<-lm(P2_log2FoldChange ~ P1_log2FoldChange, dfplotting)
#Model validation
plot(modelP)
#summary(modelP)
intercept = format(coef(modelP)[1], digits = 2)
slope = format(coef(modelP)[2], digits = 2)
rsq<-format(summary(modelP)$r.squared, digits=2)

#_______
#Plot
pdf("SpikedP1P2.pdf", height=6, width=6) 
ggplot(dfplotting) + 
  geom_point(aes(x=P1_log2FoldChange, y=P2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Pig feces", subtitle=paste("P2 = ", slope, "*P1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange P1", y="log2FoldChange P2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=P1_log2FoldChange, y=P2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")
dev.off()

P1P2List[[ "SpikedComp0hwX" ]]<-ggplot(dfplotting) + 
  geom_point(aes(x=P1_log2FoldChange, y=P2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Pig feces", x="log2FoldChange P1", y="log2FoldChange P2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=P1_log2FoldChange, y=P2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")

ggplotly(ggplot(dfplotting) + 
  geom_point(aes(x=P1_log2FoldChange, y=P2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Pig feces", subtitle=paste("P2 = ", slope, "*P1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange P1", y="log2FoldChange P2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=P1_log2FoldChange, y=P2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none"))
```

### 8 Create scatterplots sewage spiked 
```{r}
#Create df from orgs shared across all Pig fecal samples or Sewage
SharedOrgs<-merge(S1, S2, by="Genus")
#Extract log2foldchanges
dfall <- data.frame(select(SharedOrgs, matches("Genus|log2FoldChange")))
#______
#Merge for plotting
dfplotting <- merge(dfall, Feature, by="Genus")
#Create column org grouping for colouring 
dfplotting$col <- ifelse(dfplotting$Phylum=="Firmicutes", "Firmicutes",
          ifelse(dfplotting$Phylum=="Proteobacteria", "Proteobacteria",
          ifelse(dfplotting$Phylum=="Actinobacteria", "Actinobacteria",
          ifelse(dfplotting$Phylum=="Bacteroidetes", "Bacteroidetes",
          ifelse(dfplotting$Domain=="Eukaryota", "Eukaryota",
          ifelse(dfplotting$Domain=="Viruses", "Viruses",
          ifelse(dfplotting$Domain=="Archaea", "Archaea",
                 "Other")))))))
dfplotting$col[dfplotting$Class=="Negativicutes"] <- "Firmicutes(Negativicutes)"
dfplotting$col <- factor(dfplotting$col, levels=c("Proteobacteria", "Bacteroidetes", "Actinobacteria", "Firmicutes", "Firmicutes(Negativicutes)", "Eukaryota", "Archaea", "Viruses", "Other"))

#Linear model info
modelS<-lm(S2_log2FoldChange ~ S1_log2FoldChange, dfplotting)
#Model validation
plot(modelS)
#summary(modelS)
intercept = format(coef(modelS)[1], digits = 2)
slope = format(coef(modelS)[2], digits = 2)
rsq<-format(summary(modelS)$r.squared, digits=2)

#_______
#Plot
pdf("SpikedS1S2.pdf", height=6, width=6) 
ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", subtitle=paste("S2 = ", slope, "*S1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")
dev.off()

S1S2List[[ "SpikedComp0hwX" ]]<-ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none")


ggplotly(ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=1/2) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", subtitle=paste("S2 = ", slope, "*S1 + ", intercept, ", R^2 = ", rsq,    sep=""), x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="none"))
```



### 9 Create figures 
```{r}


#Have the plots stored in lists
lay <- rbind(c(2,1),
             c(4,3),
             c(6,5),
             c(8,7),
             c(10,9),
             c(12,11),
             c(12,11),
             c(14,13),
             c(14,13))
pdf(paste("SF7to9DESeqAss.pdf", sep=""), width=17, height=30)
grid.arrange(MAList$MAP1spiked, MAList$MAP1unspiked, MAList$MAP2spiked, MAList$MAP2unspiked, MAList$MAS1spiked, MAList$MAS1unspiked, MAList$MAS2spiked, MAList$MAS2unspiked, AllList$SpikedP0hvs5C64h, AllList$UnspikedComp0hwX, P1P2List$SpikedComp0hwX, P1P2List$UnspikedComp0hwX, S1S2List$SpikedComp0hwX, S1S2List$UnspikedComp0hwX, layout_matrix = lay)   
dev.off() 


#Have the plots stored in lists
lay <- rbind(c(2,2,1,1),
             c(4,4,3,3),
             c(6,6,5,5),
             c(8,8,7,7),
             c(10,10,9,9),
             c(12,14,11,13))
pdf(paste("SF7to9DESeqAssAlternative.pdf", sep=""), width=17, height=23.5)
grid.arrange(MAList$MAP1spiked, MAList$MAP1unspiked, MAList$MAP2spiked, MAList$MAP2unspiked, MAList$MAS1spiked, MAList$MAS1unspiked, MAList$MAS2spiked, MAList$MAS2unspiked, AllList$SpikedP0hvs5C64h, AllList$UnspikedComp0hwX, P1P2List$SpikedComp0hwX, P1P2List$UnspikedComp0hwX, S1S2List$SpikedComp0hwX, S1S2List$UnspikedComp0hwX, layout_matrix = lay) 
dev.off()

#Create legend and add in inkscape
legend<-ggplot(dfplotting) + 
  geom_point(aes(x=S1_log2FoldChange, y=S2_log2FoldChange, color=col, group=Genus), size=5) + 
  scale_color_manual(values=Orgcol) + 
  labs(title="Plot log2FoldChange Sewage", x="log2FoldChange S1", y="log2FoldChange S2") + 
  #geom_abline(intercept=0, slope=1) +
  geom_smooth(aes(x=S1_log2FoldChange, y=S2_log2FoldChange), method=lm) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position="bottom", legend.title=element_blank())
legendplot<-get_legend(legend)
pdf(paste("LegendSF7to9", "Genus", ".pdf", sep=""), width=17, height=12)
grid.arrange(legendplot)
dev.off()
```

## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```