
---
title: "Variables of importance (Adonis::Vegan)"
author: "Casper Sahl Poulsen"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold", 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = "center")
```

## Introduction
The purpose of the following code is to perform multivariate analysis of variances to compare storage groups using vegan::adonis, model assumptions assessed using vegan::betadisper.  

Betadisper can only evaluate one factor. Can run it on experiment and time and temperature combined.  



### Packages 
```{r}
library(vegan) #Community ecology package, ordination methods, diversity analysis and other functions for community and vegetation ecologists. help(package="vegan")
library(dplyr) #A grammar of data manipulation. A fast, consistent tool for working with data frame like objects. help(package="dplyr")
library(knitr) #A General-Purpose Package for Dynamic Report Generation in R help(package="knitr")
```

## Define variables
Define variables used in the analysis. Have to change test factor in code line 184&195 (remember to update) 
Also possible to update the number of permutations
```{r}
#Subset data pig feces 1 and 2 (P1, P2), Sewage 1 and 2 (S1, S2), or spiked unspiked
Subset<- c("Allunspiked", "PFunspiked", "SWunspiked", "P1unspiked", "P2unspiked", "S1unspiked", "S2unspiked", "Allspiked", "PFspiked", "SWspiked", "P1spiked", "P2spiked", "S1spiked", "S2spiked")
#Experiment: c("Allunspiked", "PFunspiked", "SWunspiked", "Allspiked", "PFspiked", "SWspiked") 
#TimeTemp: c("Allunspiked", "PFunspiked", "SWunspiked", "P1unspiked", "P2unspiked", "S1unspiked", "S2unspiked", "Allspiked", "PFspiked", "SWspiked", "P1spiked", "P2spiked", "S1spiked", "S2spiked")

#Subset experiment, Experiment_Type. (Meaningfull combinations of subset and SubExp: HX=All, FTX=P1&S1, LTX=P1&S1, LPSX=All unspiked, HXFTX=P1&S1, HXLTX=P1&S1, HXLPSX=All unspiked, HXLTXLPSX=P1&S1 unspiked)
SubExp<-"HX" #HX, FTX, LTX, LPSX, HXFTX, HXLTX, HXLPSX, HXLTXLPSX, HXFTXLTX, All 

#Subset Frozen Unfrozen
SubFre<-"Both" #Frozen, Unfrozen, Both

#Vegdist dissimilarity method
Dist<-"bray" #manhattan, euclidean, canberra, bray, kulczynski, jaccard, gower, altGower, morisita, horn, mountford, raup, binomial, chao, cao or mahalanobis

#Decostand standardization method
Stand<-"hellinger" #total, max, freq, normalize, range, standardize, pa, chi.square, hellinger, log

#Filtering of the count table. This is especially important to remove organisms after subsetting that are for instance not represented in pig feces 1 spiked ..., In the MAplot you can also see that the removed organisms are the low abundant ones that are already shrunk
Filter_Rows<-0 #Is based on row sums so can take integer values 
```


## Analysis
### 1 Read in data
```{r}
Tax <- read.delim(file="Taxonomy20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
Metadata <- read.delim(file="Metadata20180925.txt", check.names=FALSE, stringsAsFactors=FALSE, strip.white=TRUE)
```

### 2 Subsetting, adonis, betadisper
Creating tables for all individual tests. Additional overview of p-values created. To see evaluation plots remove hashtags.   
Using for loop to go through Subset vector
```{r}
#Make empty dataframe to add p-values
dfp<-data.frame()
dfp<-data.frame(Test=c("adonis", "betadisperAnova", "betadisperPermutest"))

for (i in Subset) {
#Removing negative and positive controls
Metadata2<-filter(Metadata, Sample_type_simple=="Sample")

#Subsetting Metadata2
if (i=="Allspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Spiked")
} else if (i=="Allunspiked") {
  Metadata2<-filter(Metadata2, SpikedUnspiked == "Unspiked")
} else if (i=="PF") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1" | Experiment == "Pig_feces_2")
} else if (i=="PFspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked" | Sample_type == "Pig_feces_2_spiked")
} else if (i=="PFunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1" | Sample_type == "Pig_feces_2")
} else if (i=="SW") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1" | Experiment == "Sewage_2")
} else if (i=="SWspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked" | Sample_type == "Sewage_2_spiked")
} else if (i=="SWunspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1" | Sample_type == "Sewage_2")
} else if (i=="P1") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_1")
} else if (i=="P1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1_spiked")
} else if (i=="P1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_1")
} else if (i=="P2") {
  Metadata2<-filter(Metadata2, Experiment == "Pig_feces_2")
} else if (i=="P2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2_spiked")
} else if (i=="P2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Pig_feces_2")
} else if (i=="S1") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_1")
} else if (i=="S1spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1_spiked")
} else if (i=="S1unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_1")
} else if (i=="S2") {
  Metadata2<-filter(Metadata2, Experiment == "Sewage_2")
} else if (i=="S2spiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2_spiked")
} else if (i=="S2unspiked") {
  Metadata2<-filter(Metadata2, Sample_type == "Sewage_2")
} else if (i=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubExp=="HX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment")
} else if (SubExp=="FTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="LTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="LPSX") {
  Metadata22<-filter(Metadata2, Experiment_type == "Library_prep_seq_platform_experiment")
  VectorLPSX <- unique(Metadata22$Matching_samples)
  Metadata2<-filter(Metadata2, Matching_samples %in% VectorLPSX)
  rm(VectorLPSX, Metadata22)
} else if (SubExp=="HXFTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment")
} else if (SubExp=="HXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="HXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXLTXLPSX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Long_term_storage_experiment" | Experiment_type == "Library_prep_seq_platform_experiment")
} else if (SubExp=="HXFTXLTX") {
  Metadata2<-filter(Metadata2, Experiment_type == "Handling_experiment" | Experiment_type == "Freeze_thaw_experiment" | Experiment_type == "Long_term_storage_experiment")
} else if (SubExp=="All") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Further subsetting Metadata2
if (SubFre=="Frozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Freezer")
} else if (SubFre=="Unfrozen") {
  Metadata2<-filter(Metadata2, FrozenUnfrozenSimple == "Unfrozen")
} else if (SubFre=="Both") {
  print("No subsetting, all included")
} else {
  print("Subset defined not valid")
}

#Applying subsetting to OTU tables 
Tax2<-dplyr::select(Tax, one_of(Metadata2$Sample))

#Filtering of the Counttable depending on rowSums. 
Tax2 <- Tax2[rowSums(Tax2)>Filter_Rows*length(Tax2),]

#Make column that combine all storage conditions to individual factors in the following test
Metadata2$TimeTemp<-paste(Metadata2$Temperature, Metadata2$Time, sep="_")

#Calculate sample-distance matrix 
distmatrix <- vegdist(decostand(t(Tax2), method=Stand), method=Dist)

#Which factor to test can pick Metadata2 columns: Experiment and TimeTemp 
Test_Factor <- "TimeTemp" #Experiment, TimeTemp
#Create design formula
design <- formula(paste("distmatrix"," ~ ", Test_Factor, sep=""))
#adonis can handle both continous and factor predictors 
set.seed(999)
adonisObject<-adonis2(design, Metadata2, by="margin", perm=9999) #, perm=999 can increase to get exact p-values  
#adonisObject #If significant then difference between groups 
#Sys.sleep(1)
write.table(data.frame(adonisObject), file=paste(i, "Adonis", "_", Test_Factor, ".txt", sep=""), sep="\t", dec=".", row.names = T, quote = F)

#### Evaluating the model assumptions  
TestModel <- with(Metadata2, betadisper(distmatrix, TimeTemp)) #Also have to change here Experiment or TimeTemp Can not run betadisper with multiple independant variables
#TestModel
#plot(TestModel)
plot(TestModel, label=FALSE)
#boxplot(TestModel)
#anova(TestModel)
anovaTestModel<-anova(TestModel) #p>0.05 -> Assumption met
write.table(data.frame(anova(TestModel)), file=paste(i, "Betadisper", "_", Test_Factor, ".txt", sep=""), sep="\t", dec=".", row.names = T, quote = F)
set.seed(999)
#permutest(TestModel)
permy<-permutest(TestModel) 
#data.frame(permy$tab)
write.table(data.frame(permy$tab), file=paste(i, "Permutest", "_", Test_Factor, ".txt", sep=""), sep="\t", dec=".", row.names = T, quote = F)
#p>0.05 -> Assumption met

TukeyHSD(TestModel, perm=999) #Increasing perm does not change p adj
dfp[paste(i)]<-c(adonisObject$`Pr(>F)`[1], anovaTestModel$`Pr(>F)`[1], permy$tab$`Pr(>F)`[1])
}

write.table(dfp, file=paste("Pvalues", Test_Factor, ".txt", sep=""), sep="\t", dec=".", row.names = T, quote = F)
```

### 3 p-values from tests
adonis if significant there is a difference between groups defined by Test_Factor  
betadisper - Anova and Permutest differences in group homogeneities are not violated if above 0.05  
```{r}
kable(dfp, caption=paste(Test_Factor))
```


## Additional
### Session information
```{r session_info}
sessionInfo()
```

### This document was processed on: 
```{r}
Sys.Date()
```